### 2023-02-03 products/manuscript/Manuscript.qmd

---
title: "Manuscript Template for a Data Analysis Project"
subtitle: "Exploring Factors influencing HIV Death-to-Case ratio in the United States and Puerto Rico -- 2020"
author: Hayley Hemme
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. 
You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(here)
library(knitr)
library(tidyverse)
library(readxl)
library(lubridate)
```

# Summary/Abstract
_Write a summary of your project._

{{< pagebreak >}}

# Introduction

## General Background Information
_Provide enough background on your topic that others can understand the why and how of your analysis_

## Description of data and data source
The data used in this analysis was retrieved from AIDSVu, a publicly available data source presented by Emory University's Rollins School of Public Health and partners, whose mission is to make HIV data widely available and accessible. AIDSVu compiled the data used in this analysis from the CDC. These datasets contain data collected at the state-level from 2020 on: new HIV diagnoses, HIV-related deaths, counts/ percentages of those with late HIV diagnoses, viral suppression, PrEP coverage, HIV prevalence, and state testing, and social determinants of health. These datasets contain demographic-level information for sex, race/ethnicity, transmission category, among others.

## Questions/Hypotheses to be addressed
Question: Can HIV mortality-prevalence ratio at the state-level in the United States be predicted by social determinant of health (ex. poverty rates), percentage of people living with HIV (PLWH) who are virally suppressed, and/ or percentage of late diagnoses/ percentage of minority HIV deaths?

https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7842853/

where mortality-prevalence ratio = (number of deaths/number of prevalant cases) / prevelance rate

H0 : Death-to-case ratio is not correlated with differences in demographic factors, social determinants of health, testing prevalence, or factors influencing HIV outcomes, such as late diagnoses, viral suppression, and PrEP coverage.

HA : Death-to-case ratio is  correlated with differences in demographic factors, social determinants of health, testing prevalence, or factors influencing HIV outcomes, such as late diagnoses, viral suppression, and PrEP coverage.

Special interest will be given to those disproportionately impacted by HIV, including Non-Hispanic Black populations and the MSM community.

To cite other work (important everywhere, but likely happens first in introduction), make sure your references are in the bibtex file specified in the YAML header above (here `dataanalysis_template_references.bib`) and have the right bibtex key. Then you can include like this:

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

_Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement._

## Data aquisition
_As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next._

## Data import and cleaning
_Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along._
Loading the datasets.
```{r}
deaths <- read_xlsx("../../data/raw_data/AIDSVu_State_Death_2020.xlsx", skip = 3)
dx <- read_xlsx("../../data/raw_data/AIDSVu_State_NewDX_2020.xlsx", skip = 3)
```

##Selecting only states/ districts in the South geographic region, as defined by the CDC.
```{r}
deaths <- deaths %>% filter(State %in% c("Alabama", "Arkansas", "Delaware", "Washington, D.C.", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia")) %>% 
  select(!contains(c("Rate Stability", "Percent", "Case")))

sex_deaths_sub <- deaths %>% select(1:7)
race_deaths_sub <- deaths  %>% select(c(1:5, 8:14))
age_deaths_sub <- deaths  %>% select(c(1:5, 15:19))

sex_deaths_sub <- sex_deaths_sub %>% pivot_longer(6:7, names_to = "Sex", values_to = "Death Rate by Sex")
sex_deaths_sub$Sex <- gsub(" Death Rate", "", sex_deaths_sub$Sex)
sex_deaths_sub <- sex_deaths_sub %>% mutate_if(is.character, as.factor) %>% 
  mutate(`GEO ID` = as.factor(`GEO ID`))

race_deaths_sub <- race_deaths_sub  %>% pivot_longer(6:12, names_to = "Race/Ethnicity", values_to = "Death Rate by Race/Ethnicity")
race_deaths_sub$`Race/Ethnicity` <- gsub(" Death Rate", "",race_deaths_sub$`Race/Ethnicity`)
race_deaths_sub <- race_deaths_sub %>% mutate_if(is.character, as.factor) %>% 
  mutate(`GEO ID` = as.factor(`GEO ID`))

age_deaths_sub <- age_deaths_sub %>% pivot_longer(6:10, names_to = "Age_Group", values_to = "Death Rate by Age Group")
age_deaths_sub$Age_Group <- gsub("Age ", "", age_deaths_sub$Age_Group)
age_deaths_sub$Age_Group <- gsub(" Death Rate", "", age_deaths_sub$Age_Group)

age_deaths_sub <- age_deaths_sub %>% mutate_if(is.character, as.factor) %>% 
  mutate(`GEO ID` = as.factor(`GEO ID`))
```

```{r}
dx <- dx %>% filter(State %in% c("Alabama", "Arkansas", "Delaware", "Washington, D.C.", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia")) %>% 
  select(!contains(c("Rate Stability", "Percent", "Case")))

sex_dx_sub <- dx %>% select(1:7)
race_dx_sub <- dx %>% select(c(1:5, 8:14))
age_dx_sub <- dx %>% select(c(1:5, 15:19))

sex_dx_sub <- sex_dx_sub %>% pivot_longer(6:7, names_to = "Sex", values_to = "New Diagnoses Rate by Sex")
sex_dx_sub$Sex <- gsub("New Diagnoses ", "", sex_dx_sub$Sex)
sex_dx_sub$Sex <- gsub(" Rate", "", sex_dx_sub$Sex)
sex_dx_sub <- sex_dx_sub %>% mutate_if(is.character, as.factor) %>% 
  mutate(`GEO ID` = as.factor(`GEO ID`))

race_dx_sub <- race_dx_sub  %>% pivot_longer(6:12, names_to = "Race/Ethnicity", values_to = "New Diagnoses Rate by Race/Ethnicity")
race_dx_sub$`Race/Ethnicity` <- gsub("New Diagnoses ", "", race_dx_sub$`Race/Ethnicity`)
race_dx_sub$`Race/Ethnicity` <- gsub(" Rate", "", race_dx_sub$`Race/Ethnicity`)
race_dx_sub <- race_dx_sub %>% mutate_if(is.character, as.factor) %>% 
  mutate(`GEO ID` = as.factor(`GEO ID`))

age_dx_sub <- age_dx_sub %>% pivot_longer(6:10, names_to = "Age_Group", values_to = "New Diagnoses Rate by Age Group")
age_dx_sub$Age_Group <- gsub("New Diagnoses Age ", "", age_dx_sub$Age_Group)
age_dx_sub$Age_Group <- gsub(" Rate", "", age_dx_sub$Age_Group)
age_dx_sub <- age_dx_sub %>% mutate_if(is.character, as.factor) %>% 
  mutate(`GEO ID` = as.factor(`GEO ID`))
```

Calculate mortality to incidence ratio (MIR) "The mortality-to-incidence ratio (MIR) can be computed from readily accessible, public-use data on cancer incidence and mortality, and a high MIR value is an indicator of poor survival relative to incidence."

The IMR (incidence:mortality ratio) is defined as the ratio of the number of people who become HIV infected per year to the number of people among those already infected who die (from any cause) per year. When this number is greater than one, the size of the population of PLHIV grows; when it is less than one, the size of that population shrinks 
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6201869/#!po=16.6667

https://www.kff.org/global-health-policy/issue-brief/assessing-global-hiv-targets-in-pepfar-countries-a-dashboard/
```{r}
sex_df <- full_join(sex_deaths_sub, sex_dx_sub)

sex_df  <- sex_df %>% group_by(State, Sex) %>%
  mutate(d2c_ratio_sex = sum(`Death Rate by Sex`/`New Diagnoses Rate by Sex`)) %>% 
  mutate(d2c_ratio_sex = round(d2c_ratio_sex, 1))

race_df <- full_join(race_deaths_sub, race_dx_sub)

race_df <- race_df %>% group_by(State, `Race/Ethnicity`) %>%
  mutate(d2c_ratio_race = sum(`Death Rate by Race/Ethnicity`/`New Diagnoses Rate by Race/Ethnicity`)) %>% 
  mutate(d2c_ratio_race = round(d2c_ratio_race, 1))

age_df <- full_join(age_deaths_sub, age_dx_sub)

age_df <- age_df %>% group_by(State, Age_Group) %>%
  mutate(d2c_ratio_age = sum(`Death Rate by Age Group`/`New Diagnoses Rate by Age Group`)) %>% 
  mutate(d2c_ratio_age = round(d2c_ratio_age, 1))

saveRDS(sex_df, "../../data/sex_df.rds")
saveRDS(race_df, "../../data/race_df.rds")
saveRDS(age_df, "../../data/age_df.rds")
```

```{r}
rm(list=ls())

sex_df <- readRDS("../../data/sex_df.rds")
race_df <- readRDS("../../data/race_df.rds")
age_df <- readRDS("../../data/age_df.rds")

```

```{r}

```

```{r}
df <- full_join(deaths, new_dx)
df <- full_join(df, prep)
df<- full_join(df, prev)

df <- df %>% select(!contains("Rate Stability"))

saveRDS(df, "../../data/raw_data/raw_df.rds")

rm(list=ls())
rm(list=ls())

prep <- read_xlsx("../../data/raw_data/AIDSVu_State_PrEP_2020.xlsx", skip = 3)
prep <- prep %>% select(!contains("Rate Stability"))

df <- readRDS("../../data/raw_data/raw_df.rds")

late_dx <- read_xlsx("../../data/raw_data/AIDSVu_State_LateDiagnoses_2020.xlsx", skip = 3)
sdoh <- read_xlsx("../../data/raw_data/AIDSVu_State_SDOH_2020.xlsx", skip = 3)
v_sup <- read_xlsx("../../data/raw_data/AIDSVu_State_ViralSuppression_2020.xlsx", skip = 3)
testing  <- read_xlsx("../../data/raw_data/AIDSVu_State_Testing_2020.xlsx", skip = 3)
```
## Statistical analysis
Explore central tendency across demographic groups. Fit a poisson regression model predicting death-to-case ratio from percentage with x SDOH variables and percentage of x demographic group.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

_Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project._

@tbl-summarytable shows a summary of the data.

Note the loading of the data providing a **relative** path using the `../../` notation. (Two dots means a folder up). You never want to specify an **absolute** path like `C:\ahandel\myproject\results\` because if you share this with someone, it won't work for them since they don't have that path. You can also use the `here` R package to create paths. See examples of that below.

```{r}
#| label: tbl-summarytable
#| tbl-cap: "Data summary table."
#| echo: FALSE
resulttable=readRDS("../../results/summarytable.rds")
knitr::kable(resulttable)
```

## Basic statistical analysis

_To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p<0.05 means statistical significance" interpretation is not valid._

@fig-result shows a scatterplot figure produced by one of the R scripts.

```{r}
#| label: fig-result
#| fig-cap: "Height and weight stratified by sex."
#| echo: FALSE
knitr::include_graphics(here("results","height_weight_stratified.png"))
```

## Full analysis

_Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here._

Example @tbl-resulttable2 shows a summary of a linear model fit.

```{r}
#| label: tbl-resulttable2
#| tbl-cap: "Linear model fit table."
#| echo: FALSE
resulttable2 = readRDS(here("results","resulttable2.rds"))
knitr::kable(resulttable2)
```

{{< pagebreak >}}

# Discussion

## Summary and Interpretation
_Summarize what you did, what you found and what it means._

## Strengths and Limitations
_Discuss what you perceive as strengths and limitations of your analysis._

## Conclusions
_What are the main take-home messages?_

_Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end_

This paper [@leek2015] discusses types of analyses.

These papers [@mckay2020; @mckay2020a] are good examples of papers published using a fully reproducible setup similar to the one shown in this template.

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References