### 2023-02-24 code/analysis_code/archieved/EDA.Rmd

```{r}
library(here)
library(GGally)
library(tidyverse)
```

```{r}
df <- readRDS(here("data/processed_data/dataframe.rds"))
df_sdoh20 <- readRDS(here("data/processed_data/df20.rds"))
```

Creating variable for HIV death rate
```{r}
df <- df %>% mutate(HIV_Rate = HIV_Deaths/Population*1e5) %>% 
  mutate(HIV_Rate = as.numeric(HIV_Rate))
df <- df %>% mutate(logHIV_Rate = log10(HIV_Rate + 1))
```

```{r}
df_DS <- df %>% group_by(Year, District, Race_Ethnicity, Sex, Age_Group) %>% 
  summarise(HIV_Deaths_DS = sum(HIV_Deaths),
            Population_DS = sum(Population),
            Uninsured_DS = mean(`% Uninsured`),
            Poverty_Rate_DS = mean(Poverty_Rate_1620)) %>% 
  mutate(HIV_Rate_DS = HIV_Deaths_DS/Population_DS*1e5)
```

```{r}
df %>% 
  group_by(District) %>% 
  drop_na(HIV_Rate) %>% 
  ggplot() +
  geom_histogram(aes(x = HIV_Rate), bins = 20)

df %>% 
  group_by(District) %>% 
  drop_na(HIV_Rate) %>% 
  ggplot() +
  geom_histogram(aes(x = logHIV_Rate))

df %>% 
  group_by(District) %>% 
  drop_na(HIV_Rate) %>%
  filter(HIV_Rate > 0) %>% 
  ggplot() +
  geom_histogram(aes(x = logHIV_Rate))
```

```{r}
x_test <- df %>%
  drop_na(HIV_Rate) %>%
  filter(HIV_Rate > 0)

shapiro.test(x_test$HIV_Rate)
```

```{r}

```