### 2023-02-24 code/analysis_code/archieved/more wrangling.Rmd

```{r}
library(here)
library(tidyverse)
```

```{r}
dths <- readRDS(here("data/processed_data/df.rds"))
dx <- readRDS(here("data/processed_data/dx.rds"))
sdoh <- readRDS(here("data/processed_data/sdoh.rds"))
pov <- readRDS(here("data/processed_data/pov.rds"))
c_hlth <- readRDS(here("data/processed_data/c_hlth.rds"))
```

Let's look into the diagnoses data a bit more. 
```{r}
glimpse(dx)
range(dx$`New Diagnoses Rate`)
```

Since -1 is not a possible value a rate can take, I referred back to AIDSVu.org and found that: '-1: Data are not shown to protect privacy because a small number of cases and/or a small population size for reasons listed in the Data Suppression and Rate Stability section.' A this moment, let's change observations with '-1' to NA.

```{r}
dx <- dx %>% mutate(`New Diagnoses Rate` = if_else(`New Diagnoses Rate` < 0, NA, `New Diagnoses Rate`))

pov <- pov %>% mutate(Race_Ethnicity = as.factor(Race_Ethnicity))

sdoh <- sdoh %>% select(!`Median Household Income`)
```

```{r}
df <- list(dths, dx, c_hlth, pov)

df <- reduce(df, full_join)

df20 <- df %>% filter(Year == 2020)

df20 <- full_join(df20, sdoh)
```

Let's save these dataframes.
```{r}
saveRDS(df20, "data/processed_data/df20.rds")
saveRDS(df, "data/processed_data/dataframe.rds")
```