### 2023-02-24 code/processing_code/archieved/1_cleaning.Rmd

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(readxl)
library(lubridate)
library(stringr)
library(tidyverse)
library(here)
```

OASIS for deaths and population (Hispanic, NH White, NH Black) for persons ages 13-59 from 2018 to 2020.

```{r warning=FALSE}
his_deaths <- read_xlsx(here("data/raw_data/hivdeaths_his_1820.xlsx"), skip = 1)
nhw_deaths <- read_xlsx(here("data/raw_data/hivdeaths_nhw_1820.xlsx"), skip = 1)
nhb_deaths <- read_xlsx(here("data/raw_data/hivdeaths_nhb_1820.xlsx"), skip = 1)

his_pop <- read_xlsx(here("data/raw_data/pop_his_1820.xlsx"), skip = 1)
nhw_pop <- read_xlsx(here("data/raw_data/pop_nhw_1820.xlsx"), skip = 1)
nhb_pop <- read_xlsx(here("data/raw_data/pop_nhb_1820.xlsx"), skip = 1)
```

Loading dataset from AIDSVu for social determinants of health (2020) and new HIV diagnoses from 2018 to 2020.
```{r warning=FALSE}
sdoh <- read_xlsx(here("data/raw_data/AIDSVu_County_SDOH_2020.xlsx"), skip = 3)

dx18 <- read_xlsx(here("data/raw_data/AIDSVu_County_NewDX_2018.xlsx"), skip = 3)
dx18 <- dx18 %>%  filter(State == "Georgia") %>% 
  select(!contains(c("Rate Stability", "Percent", "Case")))

dx19 <- read_xlsx(here("data/raw_data/AIDSVu_County_NewDX_2019.xlsx"), skip = 3)
dx19 <- dx19 %>%  filter(State == "Georgia")%>% 
  select(!contains(c("Rate Stability", "Percent", "Case")))

dx20 <- read_xlsx(here("data/raw_data/AIDSVu_County_NewDX_2020.xlsx"), skip = 3)
dx20 <- dx20 %>%  filter(State == "Georgia") %>% 
  select(!contains(c("Rate Stability", "Percent", "Case")))
```

Let's clean the `sdoh`. I'll start by filtering out all observations except those from Georgia, then I'll do a bit of reformatting.
```{r}
glimpse(sdoh)
sdoh <- sdoh %>%  filter(State == "Georgia")
sdoh$County <- str_remove(sdoh$County, " County")

sdoh <- sdoh %>% select(4:12) %>% 
  mutate(Year = "2020") %>% 
  mutate(Year = as.numeric(Year)) %>% 
  mutate(County = as.factor(County))

sdoh <- sdoh %>% select(!`Median Household Income`)
```

Let's load data from Georgia Data on Poverty rates in Georgia by race/ ethnicity for the period 2016-2020.

```{r warning = FALSE}
pov <- read_xlsx(here("data/raw_data/economics_22.xlsx"), sheet = "poverty") 
pov <- pov[c(1:159),]

glimpse(pov)
```

We'll just be using the data from the "2016-2020 Persons Below Poverty Level" by race/ethnicity. We'll also change the text formatting of County to match the other data frames.

```{r}
pov <- pov %>% select(1, starts_with("2016-2020 Persons Below Poverty Level"))

pov$County <- str_to_title(pov$County)
pov$County <- str_replace(pov$County, "Mcintosh", "McIntosh")
pov$County <- str_replace(pov$County, "Mcduffie", "McDuffie")
pov$County <- str_replace(pov$County, "Dekalb", "DeKalb")

pov <- pov %>%
  mutate(`2016-2020 Persons Below Poverty Level, Hispanic, Percent` = as.numeric(`2016-2020 Persons Below Poverty Level, Hispanic, Percent`)) %>% 
  mutate(across(where(is.numeric), function(x) x*100)) %>% 
  mutate_if(is.character, as.factor) %>% 
  pivot_longer(c(2:5), names_to = "Race_Ethnicity", values_to = "Poverty_Rate_1620")

pov$Race_Ethnicity <- str_remove(pov$Race_Ethnicity, "2016-2020 Persons Below Poverty Level, ")
pov$Race_Ethnicity <- str_remove(pov$Race_Ethnicity, ", Percent")
pov$Race_Ethnicity <- str_replace(pov$Race_Ethnicity, "Percent", "All")

pov <- pov %>% 
  filter(Race_Ethnicity %in% c("Black", "Hispanic", "White")) %>% 
  mutate(Race_Ethnicity = as.factor(Race_Ethnicity))

glimpse(pov)
```

Let's do some cleaning on the OASIS data. We'll remove extra columns and rows, as well as rows with "Selected ages/sexes total", pivot Year and Deaths into long format, and create the new variable for race/ethnicity.

```{r}
his_deaths <- his_deaths[c(-1,-2387:-2414),-7]
his_deaths <- his_deaths %>% filter(!grepl("Selected", Age)) %>% filter(!grepl("Selected", Sex)) %>%
  pivot_longer(c(4:6), names_to = "Year", values_to = "HIV_Deaths") %>% 
  mutate("Race_Ethnicity" = "Hispanic")

nhw_deaths <- nhw_deaths[c(-1,-2387:-2414),-7]
nhw_deaths <- nhw_deaths %>% filter(!grepl("Selected", Age)) %>% filter(!grepl("Selected", Sex)) %>%
  pivot_longer(c(4:6), names_to = "Year", values_to = "HIV_Deaths") %>% 
  mutate("Race_Ethnicity" = "White")

nhb_deaths <- nhb_deaths[c(-1,-2387:-2414),-7]
nhb_deaths <- nhb_deaths %>% filter(!grepl("Selected", Age)) %>% filter(!grepl("Selected", Sex)) %>%
  pivot_longer(c(4:6), names_to = "Year", values_to = "HIV_Deaths") %>% 
  mutate("Race_Ethnicity" = "Black")
```

Next, we'll merge the datasets into one dataframe, and remove the original datasets from the environment.

```{r}
deaths <- list(his_deaths, nhb_deaths, nhw_deaths)

deaths <- reduce(deaths, full_join)
rm(his_deaths, nhw_deaths, nhb_deaths)
```

Next, let's clean up the dataset for population using a similar method.

```{r}
his_pop <- his_pop[c(-1,-2387:-2412),-7]
his_pop <- his_pop %>% filter(!grepl("Selected", Age)) %>% filter(!grepl("Selected", Sex)) %>% 
  pivot_longer(c(4:6), names_to = "Year", values_to = "Population") %>% 
  mutate("Race_Ethnicity" = "Hispanic")

nhw_pop <- nhw_pop[c(-1,-2387:-2412),-7]
nhw_pop <- nhw_pop %>% filter(!grepl("Selected", Age)) %>% filter(!grepl("Selected", Sex)) %>% 
  pivot_longer(c(4:6), names_to = "Year", values_to = "Population") %>% 
  mutate("Race_Ethnicity" = "White")

nhb_pop <- nhb_pop[c(-1,-2387:-2412),-7]
nhb_pop <- nhb_pop %>% filter(!grepl("Selected", Age)) %>% filter(!grepl("Selected", Sex)) %>% 
  pivot_longer(c(4:6), names_to = "Year", values_to = "Population") %>% 
  mutate("Race_Ethnicity" = "Black")
```

Merging datasets and cleaning the original from the environment.

```{r}
pop <- list(his_pop, nhb_pop, nhw_pop)
pop <- reduce(pop, full_join)

pop$Population <- as.numeric(gsub(",","",pop$Population))
rm(his_pop, nhw_pop, nhb_pop)
```

Next, let's merge the two OASIS dataframes.

```{r}
df <- full_join(deaths, pop, by = c("Geography", "Sex", "Age", "Year", "Race_Ethnicity"))
```

Next, let's create a new variable for Public Health District, change some variable to factors, and rename the Age variable and Geography variable.

```{r}
df <- df %>% rename(County = Geography) %>% 
  mutate(District = case_when(County %in% c("Bartow", "Catoosa", "Chattooga", "Dade", "Floyd", "Gordon", "Haralson", "Paulding", "Polk", "Walker") ~ "1-1",
 County %in% c("Cherokee", "Fannin", "Gilmer", "Murray", "Pickens", "Whitfield") ~ "1-2",
 County %in% c("Banks", "Dawson", "Forsyth", "Franklin", "Habersham", "Hall", "Hart", "Lumpkin", "Rabun", "Stephens", "Towns", "Union", "White") ~ "2",
 County %in% c("Cobb", "Douglas") ~ "3-1",
 County %in% c("Fulton") ~ "3-2",
 County %in% c("Clayton") ~ "3-3",
 County %in% c("Gwinnett", "Newton", "Rockdale") ~ "3-4",
 County %in% c("DeKalb") ~ "3-5",
 County %in% c("Butts", "Carroll", "Coweta", "Fayette", "Heard", "Henry", "Lamar", "Meriwether", "Pike", "Spalding", "Troup", "Upson") ~ "4",
 County %in% c("Bleckley", "Dodge", "Johnson", "Laurens", "Montgomery", "Pulaski", "Telfair", "Treutlen", "Wheeler", "Wilcox") ~ "5-1",
 County %in% c("Baldwin", "Bibb", "Crawford", "Hancock", "Houston", "Jasper", "Jones", "Monroe", "Peach", "Putnam", "Twiggs", "Washington", "Wilkinson") ~ "5-2",
 County %in% c("Burke", "Columbia", "Emanuel", "Glascock", "Jefferson", "Jenkins", "Lincoln", "McDuffie", "Richmond", "Screven", "Taliaferro", "Warren", "Wilkes") ~ "6",
 County %in% c("Chattahoochee", "Clay", "Crisp", "Dooly", "Harris", "Macon", "Muscogee", "Marion", "Quitman", "Randolph", "Schley", "Stewart", "Sumter", "Talbot", "Taylor", "Webster") ~ "7",
 County %in% c("Ben Hill", "Berrien", "Brooks", "Cook", "Echols", "Irwin", "Lanier", "Lowndes", "Tift", "Turner") ~ "8-1",
 County %in% c("Baker", "Calhoun", "Colquitt", "Dougherty", "Decatur", "Early", "Grady", "Lee", "Miller", "Mitchell", "Seminole", "Terrell", "Thomas", "Worth") ~ "8-2",
 County %in% c("Bryan", "Camden", "Chatham", "Effingham", "Glynn", "Liberty", "Long", "McIntosh") ~ "9-1",
 County %in% c("Appling", "Atkinson", "Bacon", "Brantley", "Bulloch", "Candler", "Charlton", "Clinch", "Coffee", "Evans", "Jeff Davis", "Pierce", "Tattnall", "Toombs", "Ware", "Wayne") ~ "9-2",
 County %in% c("Barrow", "Clarke", "Elbert", "Greene", "Jackson", "Madison", "Morgan", "Oconee", "Oglethorpe", "Walton") ~ "10")) %>% 
   mutate(District = as.factor(District)) %>% 
   mutate(Sex = as.factor(Sex)) %>% 
   mutate(County = as.factor(County)) %>% 
   mutate(Race_Ethnicity = as.factor(Race_Ethnicity)) %>% 
   rename(Age_Group = Age) %>% 
   mutate(Age_Group = as.factor(Age_Group))
```

Let's clean up the Age_Group variable by removing text and make it a factor with levels. Changing from class from character to numeric for Year, HIV_Deaths, and Population.

```{r}
df$Age_Group <- str_replace(df$Age_Group, " Adolescence", "")
df$Age_Group <- str_replace(df$Age_Group, " Early Adulthood", "")
df$Age_Group <- str_replace(df$Age_Group, " Young Adulthood", "")
df$Age_Group <- str_replace(df$Age_Group, " Middle Adulthood", "")

df <- df %>% mutate(Age_Group = as.factor(Age_Group),
                   Age_Group = factor(Age_Group,
    level = c("13-19", "20-29", "30-44", "45-59"))) %>% 
  mutate_if(is.character, as.numeric) %>% 
  relocate(where(is.numeric), .after = last_col()) %>% 
  relocate(`District`, .after = `County`) %>% 
  relocate(`Age_Group`, .after = `Race_Ethnicity`)
glimpse(df)
```

```{r}
glimpse(df)
```
Things are looking good so far!

Now let's move onto the AIDSVu data. First, we'll merge all new HIV diagnoses from 2018-2020 into one dataframe

```{r}
dx <- list(dx18, dx19, dx20)
dx <- reduce(dx, full_join)
rm(dx18, dx19, dx20)
```

Let's remove columns for race/ethnicities other than Hispanic, NH White, and NH Black, as well as some other variables we won't be using in this analysis. We will also rename county and create the district variable.

```{r}
dx <- dx %>% select(!contains(c("Asian", "American", "Multiracial", "Hawaiian", "MSM", "2013", "GEO ID", "State")))%>% 
  rename(County = "County Name")

dx$County <- str_remove(dx$County, " County")

dx <-  dx %>% mutate(District = case_when(County %in% c("Bartow", "Catoosa", "Chattooga", "Dade", "Floyd", "Gordon", "Haralson", "Paulding", "Polk", "Walker") ~ "1-1",
 County %in% c("Cherokee", "Fannin", "Gilmer", "Murray", "Pickens", "Whitfield") ~ "1-2",
 County %in% c("Banks", "Dawson", "Forsyth", "Franklin", "Habersham", "Hall", "Hart", "Lumpkin", "Rabun", "Stephens", "Towns", "Union", "White") ~ "2",
 County %in% c("Cobb", "Douglas") ~ "3-1",
 County %in% c("Fulton") ~ "3-2",
 County %in% c("Clayton") ~ "3-3",
 County %in% c("Gwinnett", "Newton", "Rockdale") ~ "3-4",
 County %in% c("DeKalb") ~ "3-5",
 County %in% c("Butts", "Carroll", "Coweta", "Fayette", "Heard", "Henry", "Lamar", "Meriwether", "Pike", "Spalding", "Troup", "Upson") ~ "4",
 County %in% c("Bleckley", "Dodge", "Johnson", "Laurens", "Montgomery", "Pulaski", "Telfair", "Treutlen", "Wheeler", "Wilcox") ~ "5-1",
 County %in% c("Baldwin", "Bibb", "Crawford", "Hancock", "Houston", "Jasper", "Jones", "Monroe", "Peach", "Putnam", "Twiggs", "Washington", "Wilkinson") ~ "5-2",
 County %in% c("Burke", "Columbia", "Emanuel", "Glascock", "Jefferson", "Jenkins", "Lincoln", "McDuffie", "Richmond", "Screven", "Taliaferro", "Warren", "Wilkes") ~ "6",
 County %in% c("Chattahoochee", "Clay", "Crisp", "Dooly", "Harris", "Macon", "Muscogee", "Marion", "Quitman", "Randolph", "Schley", "Stewart", "Sumter", "Talbot", "Taylor", "Webster") ~ "7",
 County %in% c("Ben Hill", "Berrien", "Brooks", "Cook", "Echols", "Irwin", "Lanier", "Lowndes", "Tift", "Turner") ~ "8-1",
 County %in% c("Baker", "Calhoun", "Colquitt", "Dougherty", "Decatur", "Early", "Grady", "Lee", "Miller", "Mitchell", "Seminole", "Terrell", "Thomas", "Worth") ~ "8-2",
 County %in% c("Bryan", "Camden", "Chatham", "Effingham", "Glynn", "Liberty", "Long", "McIntosh") ~ "9-1",
 County %in% c("Appling", "Atkinson", "Bacon", "Brantley", "Bulloch", "Candler", "Charlton", "Clinch", "Coffee", "Evans", "Jeff Davis", "Pierce", "Tattnall", "Toombs", "Ware", "Wayne") ~ "9-2",
 County %in% c("Barrow", "Clarke", "Elbert", "Greene", "Jackson", "Madison", "Morgan", "Oconee", "Oglethorpe", "Walton") ~ "10")) %>% 
  mutate(District = as.factor(District)) %>% 
  mutate(County = as.factor(County))

range(dx$`New Diagnoses Rate`)
```
Since -1 is not a possible value a rate can take, I referred back to AIDSVu.org and found that: '-1: Data are not shown to protect privacy because a small number of cases and/or a small population size for reasons listed in the Data Suppression and Rate Stability section.' A this moment, let's change observations with '-1' to NA. 
```{r}
dx <- dx %>% select(County, District, Year, `New Diagnoses Rate`)%>% 
  mutate(`New Diagnoses Rate` = if_else(`New Diagnoses Rate` < 0, NA, `New Diagnoses Rate`))

glimpse(dx)
```

Here, I'm loading data from County Health rankings for years 2018 - 2020
```{r}
cnty_hlth18A <- read_xls(path = here("data/raw_data/2018 County Health Rankings Georgia Data - v3.xls"), sheet = 4, skip = 1)
cnty_hlth18B <- read_xls(path = here("data/raw_data/2018 County Health Rankings Georgia Data - v3.xls"), sheet = 5, skip = 1)
cnty_hlth18A <- cnty_hlth18A %>%  select(c(3, 67:75))
cnty_hlth18B <- cnty_hlth18B %>%  select(c(3, 62:64))
c_hlt18 <- full_join(cnty_hlth18A, cnty_hlth18B)

c_hlt18 <- c_hlt18 %>% mutate(Year = as.numeric(2018)) %>%
  select(1:3,7:9,11, 14) %>% 
  rename(`Median Household Income` = `Household Income`)
c_hlt18 <-  c_hlt18[-1,]

cnty_hlth19A <- read_xls(path = here("data/raw_data/2019 County Health Rankings Georgia Data - v1_0.xls"), sheet = 4, skip = 1)
cnty_hlth19B <- read_xls(path = here("data/raw_data/2019 County Health Rankings Georgia Data - v1_0.xls"), sheet = 5, skip = 1)
cnty_hlth19A <- cnty_hlth19A %>%  select(c(3, 67:75))
cnty_hlth19B <- cnty_hlth19B %>%  select(c(3, 66:68))
c_hlt19 <- full_join(cnty_hlth19A, cnty_hlth19B)

c_hlt19 <- c_hlt19 %>% mutate(Year = as.numeric(2019)) %>%
  select(1:3,7:9,11, 14)  %>% 
  rename(`Median Household Income` = `Household Income`)
c_hlt19 <-  c_hlt19[-1,]

cnty_hlth20A <- read_xlsx(path = here("data/raw_data/2020 County Health Rankings Georgia Data - v1_0.xlsx"), sheet = 4, skip = 1)
cnty_hlth20B <- read_xlsx(path = here("data/raw_data/2020 County Health Rankings Georgia Data - v1_0.xlsx"), sheet = 5, skip = 1)
cnty_hlth20A <- cnty_hlth20A %>%  select(c(3, 104:112))
cnty_hlth20B <- cnty_hlth20B %>%  select(c(3, 158:160))

c_hlt20 <- full_join(cnty_hlth20A, cnty_hlth20B)
c_hlt20 <- c_hlt20 %>% mutate(Year = as.numeric(2020)) %>%
  select(1:3,7:9,11, 14)    %>% 
  rename(`PCP Rate` = `Primary Care Physicians Rate`) %>% 
  rename(`PCP Ratio` = `Primary Care Physicians Ratio`)
c_hlt20 <-  c_hlt20[-1,]
```

```{r}
c_hlth <- list(c_hlt18,c_hlt19, c_hlt20)

c_hlth  <- reduce(c_hlth, full_join)
c_hlth <- c_hlth %>% mutate(County = as.factor(County))
```

Things are looking pretty good so far, so let's save these dataframes.
```{r}
saveRDS(df, here("data/processed_data/df.rds"))
saveRDS(dx, here("data/processed_data/dx.rds"))
saveRDS(sdoh, here("data/processed_data/sdoh.rds"))
saveRDS(pov, here("data/processed_data/pov.rds"))
saveRDS(c_hlth, here("data/processed_data/c_hlth.rds"))
```

Now, let's merge the datasets - dths, dx, c_hlth, pov. Since `sdoh` contains only data from 2020, we'll create another subset containing just observations from 2020 and merge with `sdoh`.
```{r}
df1 <- list(df, dx, c_hlth, pov)

df1 <- reduce(df1, full_join)

df20 <- df %>% filter(Year == 2020)

df20 <- full_join(df20, sdoh)
```

Let's save these dataframes.
```{r}
saveRDS(df20, here("data/processed_data/df20.rds"))
saveRDS(df1, here("data/processed_data/dataframe.rds"))
```