### 2023-02-24 products/manuscript/MANU.qmd

---
title: "Manuscript Template for a Data Analysis Project"
subtitle: "Exploring Factors influencing HIV-related Mortality Rates in the Georgia -- 2018 - 2020"
author: Hayley Hemme
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
bibliography: ../dataanalysis_template_references.bib
csl: ../apa.csl
---

The structure below is one possible setup for a data analysis project (including the course project). For a manuscript, adjust as needed. You don't need to have exactly these sections, but the content covering those sections should be addressed.

This uses MS Word as output format. [See here](https://quarto.org/docs/output-formats/ms-word.html) for more information. You can switch to other formats, like html or pdf. See [the Quarto documentation](https://quarto.org/) for other formats.

```{r, echo=FALSE, message=FALSE}
# load a few R packages
library(knitr)
library(readxl)
library(lubridate)
library(stringr)
library(tidyverse)
library(here)
```

# Summary/Abstract

*Write a summary of your project.*

{{< pagebreak >}}

# Introduction

## General Background Information

*Provide enough background on your topic that others can understand the why and how of your analysis*

## Description of data and data source

The data used in this project was obtained from multiple publicly available source. Data pertaining to HIV deaths, new diagnoses, HIV-related outcomes, and social determinants of health relating to HIV were obtained from AIDSVu. [AIDSVu](https://aidsvu.org/resources/#/datasets). Population data for on persons of race/ethnicities included in the analysis and total population was obtained from the [U.S. Census Bureau](https://data.census.gov/). Percentage population make-up by sex was obtained from [KFF](https://www.kff.org/other/state-indicator/) . Percentage distribution of persons under 14 was obtained from Kids Count [Kids Count](https://datacenter.kidscount.org/). 
All data described was aggregated at the the state-level.

## Questions/Hypotheses to be addressed

Question: Can the HIV death rate from 2020 in the United States be modeled by social determinants of health like poverty and/ or by HIV related outcomes (new diagnoses, prevalence PrEP use, etc?) Explain in separate analysis: Race/ ethnicity (Black, Hispanic, White, Multiple) and Sex (Male, Female). Can a spatial model explain variation in HIV death-rates at the state-level and/or regional-level after accounting for variables associated with sociodemographic, SDOH, and HIV-related outcomes?

H0 : There is not a spatial correlation HIV-related deaths and SDOH or HIV-related outcomes.
Ha : There is a spatial correlation between HIV-related deaths and SDOH or HIV-related outcomes.

Examples of reproducible research projects can for instance be found in [@mckay2020; @mckay2020a]

{{< pagebreak >}}

# Methods

*Describe your methods. That should describe the data, the cleaning processes, and the analysis approaches. You might want to provide a shorter description here and all the details in the supplement.*

## Data aquisition

*As applicable, explain where and how you got the data. If you directly import the data from an online source, you can combine this section with the next.*

## Data import and cleaning

*Write code that reads in the file and cleans it so it's ready for analysis. Since this will be fairly long code for most datasets, it might be a good idea to have it in one or several R scripts. If that is the case, explain here briefly what kind of cleaning/processing you do, and provide more details and well documented code somewhere (e.g. as supplement in a paper). All materials, including files that contain code, should be commented well so everyone can follow along.* Loading the datasets from

Clear the environment and start fresh.

```{r}

```

## Statistical analysis

Explore central tendency across demographic groups. Fit a poisson regression model predicting death-to-case ratio from percentage with x SDOH variables and percentage of x demographic group.

{{< pagebreak >}}

# Results

## Exploratory/Descriptive analysis

*Use a combination of text/tables/figures to explore and describe your data. Show the most important descriptive results here. Additional ones should go in the supplement. Even more can be in the R and Quarto files that are part of your project.*

## Basic statistical analysis

*To get some further insight into your data, if reasonable you could compute simple statistics (e.g. simple models with 1 predictor) to look for associations between your outcome(s) and each individual predictor variable. Though note that unless you pre-specified the outcome and main exposure, any "p\<0.05 means statistical significance" interpretation is not valid.*

## Full analysis

*Use one or several suitable statistical/machine learning methods to analyze your data and to produce meaningful figures, tables, etc. This might again be code that is best placed in one or several separate R scripts that need to be well documented. You want the code to produce figures and data ready for display as tables, and save those. Then you load them here.*

{{< pagebreak >}}

# Discussion

## Summary and Interpretation

*Summarize what you did, what you found and what it means.*

## Strengths and Limitations

*Discuss what you perceive as strengths and limitations of your analysis.*

## Conclusions

*What are the main take-home messages?*

*Include citations in your Rmd file using bibtex, the list of references will automatically be placed at the end*

Note that this cited reference will show up at the end of the document, the reference formatting is determined by the CSL file specified in the YAML header. Many more style files for almost any journal [are available](https://www.zotero.org/styles). You also specify the location of your bibtex reference file in the YAML. You can call your reference file anything you like, I just used the generic word `references.bib` but giving it a more descriptive name is probably better.

{{< pagebreak >}}

# References