### 2023-02-24 code/processing_code/1_cleaning_2020.Rmd
more cleaning

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(readxl)
library(lubridate)
library(stringr)
library(tidyverse)
library(here)
```

```{r}
deaths <- read_xlsx(here("data/raw_data/AIDSVu_State_Death_2020.xlsx"), skip = 3)
#late_dx <- read_xlsx(here("data/raw_data/AIDSVu_State_LateDiagnoses_2020.xlsx"), skip = 3)
us_race_pop <- read_xlsx(here("data/raw_data/DECENNIALPL2020.P2-2023-02-24T124235.xlsx"), sheet = 2)
us_sex_pop <-read_csv(here("data/raw_data/raw_data.csv"), skip = 2)

#new_dx <- read_xlsx(here("data/raw_data/AIDSVu_State_NewDX_2020.xlsx"), skip = 3)
#prev  <- read_xlsx(here("data/raw_data/AIDSVu_State_Prev_2020.xlsx"), skip = 3)
#rec_care  <- read_xlsx(here("data/raw_data/AIDSVu_State_ReceiptofCare_2020.xlsx"), skip = 3)
#sdoh  <- read_xlsx(here("data/raw_data/AIDSVu_State_SDOH_2020.xlsx"), skip = 3)
#v_sup  <- read_xlsx(here("data/raw_data/AIDSVu_State_ViralSuppression_2020.xlsx"), skip = 3)
```

Let's start by separating `deaths` by demographic characteristic . While this data provides information on demographic characteristics such as sex, age group, and race/ethnicity, it does not provide data on all of these terms together. We'll separating by sex and race/ethnicity. This analysis will be limited to Non-Hispanic Black, Non-Hispanic White, and Hispanic persons.
```{r}
deaths_sex <- deaths %>% select(1:4, contains(c("male", "MSM")))

deaths_race <-   deaths %>% select(1:4, contains(c("Black", "White", "Hispanic",  "Multiple")))

summary(deaths_race)
```

Some rates are -9 due to missingness. Let's convert these to NA
```{r}
death_race <- deaths_race %>% 
  mutate(across(ends_with("Death Rate"), ~ case_when( .x < 0 ~ NA, TRUE ~ .x)))
```

Next, let's move on to the census data. I'll start by re-formatting the data from wide format for States to long, and format race/ethnicity into wide.
```{r}
glimpse(us_race_pop)

race_pop <- us_race_pop %>%  
  pivot_longer(2:54 , names_to = "Location", values_to = "Population" ) %>%  
  pivot_wider(names_from = Label, values_from= Population)
```

```{r}
race_pop <- race_pop %>% 
  select(c(Location, starts_with(c("Total:", "Black or African American alone", "Hispanic", "White alone", "Population of two or more races:")))) %>% 
  rename(c(Hispanic = `Hispanic or Latino`,
         White = `White alone`,
         Black = `Black or African American alone`,
         Multiple = `Population of two or more races:`))
```

I need to remove all commas from the dataframe and convert the columns to numeric. 
```{r}
race_pop <- race_pop %>% 
  mutate_all(~ str_remove_all(., ",")) %>% 
  mutate(across(2:6, ~ as.numeric(.)))

```

Then, I'll calculate the counts of males and females in each location.
```{r}
total_pop <- race_pop %>% select(1:2)
pop_sex <- inner_join(total_pop, us_sex_pop)
```

Next, we need to estimate the population older than 12. I will try to estimate this using data from Kids Count.

```{r}
kid_pop <- read_xlsx(here("data/raw_data/Child population by age group.xlsx"))
kid_pop <- kid_pop %>% 
  filter(TimeFrame == 2020) %>% 
  filter(`Age group` %in% c("0 to 4", "5 to 11", "12 to 14")) %>% 
  mutate(Data = as.numeric(Data))

pop_less13 <- kid_pop %>% 
  group_by(Location) %>% 
  mutate(`Age group` = ifelse(`Age group` == '12 to 14', '12', `Age group`)) %>%
  group_by(`Age group`) %>%
  mutate(`Data` = ifelse(`Age group` == '12', `Data` * 0.333, `Data`)) %>%
  group_by(Location) %>% 
  summarise(Less13 = sum(Data))

new_pop <- full_join(pop_sex, pop_less13, by = join_by(Location))

new_pop <- new_pop %>% 
  mutate(Male_Pop = Male * `Total:`,
  Female_Pop = Female * `Total:`,
  Female_kid_pop = Female * Less13,
  Male_kid_pop = Male * Less13)

df_pop <- new_pop %>% 
  group_by(Location) %>% 
  summarize(df_pop_male = sum(Male_Pop - Male_kid_pop),
            df_pop_female = sum(Female_Pop - Female_kid_pop))

```

```{r}
sex_df <- inner_join(deaths_sex, df_pop, by = join_by(State == Location))
```

```{r}
sex_df %>% mutate(M_Rate = `Male Death Cases`/df_pop_male*1e5)
```

Next find expected rate vs observed rate