### 2023-02-25 code/processing_code/2_cleaning_sex_df.Rmd

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(here)
```

```{r}
sex_df <- readRDS(here("data/processed_data/sex_df.rds"))
```
Let's find the HIV death rate by sex
```{r}
sex_df <- sex_df %>% 
  mutate(M_Rate = round(`Male Death Cases`/pop_male*1e5),
         F_Rate = round(`Female Death Cases`/pop_female*1e5))
```

Next, let's find the expected number of HIV deaths. First, we'll calculate lambda.
```{r}
lambda_male <- sex_df %>% drop_na(`Male Death Cases`) %>% 
  drop_na(pop_male) %>% 
  summarize(lambda_male = sum(`Male Death Cases`)/sum(pop_male)) %>%
  as.numeric()

lambda_female <- sex_df %>% drop_na(`Female Death Cases`) %>% 
  drop_na(pop_female) %>% 
  summarize(lambda_female = sum(`Female Death Cases`)/sum(pop_female)) %>%
  as.numeric()

sex_df <- sex_df %>%
  mutate(male_exp_deaths = round(pop_male * lambda_male),
         female_exp_deaths = round(pop_female *  lambda_female))
```

```{r}
sex_df <- sex_df %>% 
  mutate(ratio_OE_male = `Male Death Cases` / male_exp_deaths,
         ratio_OE_female = `Female Death Cases` / female_exp_deaths)
```