### 2023-02-25 3_cleaning_race_df.Rmd
completed cleaning

```{r, echo=FALSE, message=FALSE}
library(knitr)
library(tidyverse)
library(here)
```

```{r}
race_df <- readRDS(here("data/processed_data/race_df.rds"))
```
Let's find the HIV death rate by race
```{r}
race_df <- race_df %>% 
  mutate(black_death_rate = round(`Black Death Cases`/Black_pop*1e5),
         hispanic_death_rate = round(`Hispanic Death Cases`/Hispanic_pop*1e5),
         white_death_rate = round(`White Death Cases`/White_pop*1e5),
         multiple_death_rate = round(`Multiple Race Death Cases`/Multiple_pop*1e5))
```

Next, let's find the expected number of HIV deaths. First, we'll calculate lambda.
```{r}
lambda_black <- race_df %>%
  drop_na(`Black Death Cases`) %>% 
  summarize(lambda_black = sum(`Black Death Cases`)/sum(Black_pop)) %>%
  as.numeric()

lambda_hispanic <- race_df %>%
  drop_na(`Hispanic Death Cases`) %>% 
  summarize(lambda_hispanic = sum(`Hispanic Death Cases`)/sum(Hispanic_pop)) %>%
  as.numeric()

lambda_white <- race_df %>%
  drop_na(`White Death Cases`) %>% 
  summarize(lambda_white = sum(`White Death Cases`)/sum(White_pop)) %>%
  as.numeric()

lambda_multiple <- race_df %>%
  drop_na(`Multiple Race Death Cases`) %>% 
  summarize(lambda_multipe = sum(`Multiple Race Death Cases`)/sum(Multiple_pop)) %>%
  as.numeric()

race_df <- race_df %>%
  mutate(black_exp_deaths = round(Black_pop * lambda_black),
        white_exp_deaths = round(White_pop * lambda_white),
        hispanic_exp_deaths = round(Hispanic_pop *  lambda_hispanic),
        multiple_exp_deaths = round(Multiple_pop *  lambda_multiple))
```

```{r}
race_df <- race_df %>% 
  mutate(ratio_OE_black = `Black Death Cases` / black_exp_deaths,
         ratio_OE_hispanic = `Hispanic Death Cases` / hispanic_exp_deaths,
         ratio_OE_white = `White Death Cases` / white_exp_deaths,
         ratio_OE_multiple = `Multiple Race Death Cases` / multiple_exp_deaths)
```

```{r}
black_df <- race_df %>% select(1:3, contains("Black"))

black_df <- black_df %>% 
  mutate(Race = "Black") %>% 
 rename("Death Rate (AIDSVu)" = `Black Death Rate`,
        "Population" = Black_pop,
        "Death Cases" = "Black Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Black Rate",
        "New Diagnoses Cases" = "New Diagnoses Black Cases",
        "Percentage Diagnosed Late" = `Percent of Black Persons Diagnosed Late`,
        "Number of Persons Diagnosed Late" = `Number of Black Persons Diagnosed Late`,
        "Percentage Received Care" = `Percent of Black Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of Black Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "Black Rate",
        "Prevalent HIV Cases" = "Black Cases",
        "Expected Death Cases" = black_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_black,
        "Death Rate (estimated)" = black_death_rate)

hispanic_df <- race_df %>% select(1:3, contains("Hispanic"))

hispanic_df <- hispanic_df %>% 
  mutate(Race = "Hispanic") %>% 
 rename("Death Rate (AIDSVu)" = `Hispanic Death Rate`,
        "Population" = Hispanic_pop,
        "Death Cases" = "Hispanic Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Hispanic Rate",
        "New Diagnoses Cases" = "New Diagnoses Hispanic Cases",
        "Percentage Diagnosed Late" = `Percent of Hispanic Persons Diagnosed Late`,
        "Number of Persons Diagnosed Late" = `Number of Hispanic Persons Diagnosed Late`,
        "Percentage Received Care" = `Percent of Hispanic Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of Hispanic Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "Hispanic Rate",
        "Prevalent HIV Cases" = "Hispanic Cases",
        "Expected Death Cases" = hispanic_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_hispanic,
        "Death Rate (estimated)" = hispanic_death_rate)

white_df <- race_df %>% select(1:3, contains("White"))

white_df <- white_df %>% 
  mutate(Race = "White") %>% 
 rename("Death Rate (AIDSVu)" = `White Death Rate`,
        "Population" = White_pop,
        "Death Cases" = "White Death Cases",
        "New Diagnoses Rate" = "New Diagnoses White Rate",
        "New Diagnoses Cases" = "New Diagnoses White Cases",
        "Percentage Diagnosed Late" = `Percent of White Persons Diagnosed Late`,
        "Number of Persons Diagnosed Late" = `Number of White Persons Diagnosed Late`,
        "Percentage Received Care" = `Percent of White Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of White Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "White Rate",
        "Prevalent HIV Cases" = "White Cases",
        "Expected Death Cases" = white_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_white,
        "Death Rate (estimated)" = white_death_rate)

multiple_df <- race_df %>% select(1:3, contains("Multiple"))

multiple_df <- multiple_df %>% 
  mutate(Race = "Multiple") %>% 
 rename("Death Rate (AIDSVu)" = "Multiple Race Death Rate",
        "Population" = Multiple_pop,
        "Death Cases" = "Multiple Race Death Cases",
        "New Diagnoses Rate" = "New Diagnoses Multiple Race Rate",
        "New Diagnoses Cases" = "New Diagnoses Multiple Race Cases",
        "Percentage Received Care" = `Percent of Multiple Race Persons Living with HIV who Received HIV Care`,
        "Percentage Virally Suppressed" = `Percent of Multiple Race Persons Living with HIV who were Virally Suppressed`,
        "HIV Prevalence Rate" = "Multiple Races Rate",
        "Prevalent HIV Cases" = "Multiple Races Cases",
        "Expected Death Cases" = multiple_exp_deaths,
        "Ratio Observed/Expected" = ratio_OE_multiple,
        "Death Rate (estimated)" = multiple_death_rate)

race_df <- list(black_df, hispanic_df, white_df, multiple_df)

race_df <- reduce(race_df, full_join)
```

```{r}
saveRDS(race_df, here("data/processed_data/race_df_cleaned.rds"))
```