### 2023-02-26 code/analysis_code/2_EDA_sex.Rmd

```{r}
library(here)
library(tidyverse)
library(spdep)
```
https://github.com/ubikuity/List-of-neighboring-states-for-each-US-state
List-of-neighboring-states-for-each-US-state
```{r}
sex_df <- readRDS(here("data/processed_data/sex_new.rds"))
us_nabs <- read_csv(here("data/raw_data/neighbors-states.csv"))

us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))
```

```{r}
male_df <- sex_df %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State),
         Sex == "Male")

h_male <- matrix(0, nrow(male_df), nrow(male_df))
h_male[cbind(match(us_nabs$StateCode, male_df$`State Abbreviation`), match(us_nabs$NeighborStateCode, male_df$`State Abbreviation`))]<-1
nablist_male <- mat2listw(h_male)
```

```{r}
par(mfrow=c(1,2))
moran.test(male_df$BOX, nablist_male, zero.policy = TRUE, randomisation = FALSE)
moran.plot(male_df$BOX, nablist_male, zero.policy = TRUE)

lisa_male <-localmoran(male_df$`Ratio Observed/Expected`,  nablist_male)
summary(lisa_male[,1])
```

Since the data frame for female is missing observations (due to 0) for Vermont and New Hampshire, we need to remove these from the neighbor list.
```{r}
female_df <- sex_df %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State),
         Sex == "Female")
us_nabs_f <- us_nabs %>%
  filter(!grepl("VT", StateCode),
         !grepl("NH", StateCode))

h_female <- matrix(0, nrow(female_df), nrow(female_df))
h_female[cbind(match(us_nabs_f$StateCode, female_df$`State Abbreviation`), match(us_nabs_f$NeighborStateCode, female_df$`State Abbreviation`))]<-1
nablist_female <- mat2listw(h_female)

par(mfrow=c(1,2))
moran.test(female_df$BOX, nablist_female, zero.policy = TRUE, randomisation = FALSE)
moran.plot(female_df$BOX, nablist_female, zero.policy = TRUE)

lisa_female <-localmoran(female_df$`Ratio Observed/Expected`,  nablist_female)
summary(lisa_female[,1])
```