### 2023-03-18 code/analysis_code/Race/p1_EDA_race.Rmd
more EDA and modeling

```{r message=FALSE, warning=FALSE}
library(here)
library(spdep)
library(GGally)
library(skimr)
library(ggfortify)
library(MASS)
library(ggcorrplot)
library("FactoMineR")
library(tidyverse)
```
https://github.com/ubikuity/List-of-neighboring-states-for-each-US-state
List-of-neighboring-states-for-each-US-state
```{r}
race_df <- readRDS(here("data/processed_data/race_df_cleaned.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subsetting to only contiguous states.
race_df <- race_df %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))

us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))
```

```{r}
autoplot(lm(race_df$`Death Rate (AIDSVu)` ~ 1))
autoplot(lm(race_df$`Death Rate (estimated)` ~ 1))
autoplot(lm(race_df$`Ratio Observed/Expected` ~ 1))
```

```{r}
race_df %>% 
  ggplot() +
  geom_histogram(aes(x = `Death Rate (estimated)`, fill = Race), bins = 10)
shapiro.test(race_df$`Death Rate (estimated)`)

race_df %>% 
  ggplot() +
  geom_histogram(aes(x = `Ratio Observed/Expected`, fill = Race), bins = 10)
shapiro.test(race_df$`Ratio Observed/Expected`)
```
Both histograms show a positively skewed distribution. ROE appears to be the slightest bit more normal.

```{r}
race_df %>% 
  ggplot() +
  geom_histogram(aes(x = log(`Death Rate (estimated)` + 1), fill = Race) , bins = 10)
shapiro.test(log(race_df$`Death Rate (estimated)` + 1))

race_df %>% 
  ggplot() +
  geom_histogram(aes(x = log(`Ratio Observed/Expected` + 1), fill = Race) , bins = 10)
shapiro.test(log(race_df$`Ratio Observed/Expected` + 1))
```
The log transformation of Death Rate better fit, but there is still sufficient evidence to reject normality.

```{r}
race_df_DR <- race_df %>% 
  filter(`Death Rate (estimated)` > 0)
b_DR <-  boxcox(lm(race_df_DR$`Death Rate (estimated)` ~ 1))
lambda_DR <- b_DR$x[which.max(b_DR$y)]

race_df_ROE <- race_df %>% 
  filter(`Ratio Observed/Expected` > 0)

b_ROE <-  boxcox(lm(race_df_ROE$`Ratio Observed/Expected` ~ 1))
lambda_ROE <- b_ROE$x[which.max(b_ROE$y)]
```

We'll create a function to carry out the box cox transformation [cite](https://stats.stackexchange.com/questions/1444/how-should-i-transform-non-negative-data-including-zeros)
```{r}
box_cox <- function(x, lambda) {

    eps <- 0.00001
    if (abs(lambda) < eps)
        log(x)
    else
        (x ^ lambda - 1) / lambda

}
```

Box Cox transformation of DR and results of shapiro test.
```{r}
race_df_DR %>%
  mutate(BOX = box_cox(`Death Rate (estimated)`, lambda_DR)) %>%
  ggplot() + geom_histogram(aes(x= BOX, fill = Race), bins = 10) + facet_wrap(. ~ Race)

race_df_DR <- race_df_DR %>%
  mutate(BOX = box_cox(`Death Rate (estimated)`, lambda_DR))

race_df_DR %>%
  pull(BOX) %>% 
  shapiro.test()
##Reject normality

race_df_DR %>%
  filter(Race == "Black") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Fail to reject normality

race_df_DR %>%
  filter(Race == "White") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Reject normality

race_df_DR %>%
  filter(Race == "Hispanic") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Fail to reject normality

race_df_DR %>%
  filter(Race == "Multiple") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Fail to reject normality

par(mfrow=c(1,3))
plot(lm(race_df_DR$BOX ~ 1))
```
Successful transformation of death rate among of Black, Hispanic, and Multiple racial backgrounds persons.

Box Cox transformation of ROE and results of Shapiro test.
```{r}
race_df_ROE %>%
  mutate(BOX = box_cox(`Ratio Observed/Expected`, lambda_ROE)) %>%
  ggplot() + geom_histogram(aes(x= BOX, fill = Race), bins = 10) + facet_wrap(. ~ Race)

race_df_ROE <- race_df_ROE %>%
  mutate(BOX = box_cox(`Ratio Observed/Expected`, lambda_ROE))

race_df_ROE %>%
  pull(BOX) %>% 
  shapiro.test()

race_df_ROE %>%
  filter(Race == "Black") %>% 
  pull(BOX) %>% 
  shapiro.test()

race_df_ROE %>%
  filter(Race == "White") %>% 
  pull(BOX) %>% 
  shapiro.test()

race_df_ROE %>%
  filter(Race == "Hispanic") %>% 
  pull(BOX) %>% 
  shapiro.test()

race_df_ROE %>%
  filter(Race == "Multiple") %>% 
  pull(BOX) %>% 
  shapiro.test()

par(mfrow=c(1,3))
plot(lm(race_df_ROE$BOX ~ 1))
```

Let's save the dataframe for ROE since after the Box Cox transformation for all races was better with ROE than DR.
```{r}
saveRDS(race_df_ROE, here("data/processed_data/race_new.rds"))
```