### 2023-04-08 code/analysis_code/Race/p1_EDA_race.Rmd
added more to project

```{r message=FALSE, warning=FALSE}
library(here)
library(spdep)
library(GGally)
library(skimr)
library(ggfortify)
library(MASS)
library(ggcorrplot)
library("FactoMineR")
library(tidyverse)
```
https://github.com/ubikuity/List-of-neighboring-states-for-each-US-state
List-of-neighboring-states-for-each-US-state
```{r}
race_df <- readRDS(here("data/processed_data/race_df_cleaned.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subsetting to only contiguous states.
race_df <- race_df %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))

us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))
```

```{r}
autoplot(lm(race_df$`Death Rate (AIDSVu)` ~ 1))
autoplot(lm(race_df$`Death Rate (estimated)` ~ 1))
autoplot(lm(race_df$`Ratio Observed/ Expected` ~ 1))
```

```{r}
race_df %>% 
  ggplot() +
  geom_histogram(aes(x = `Death Rate (estimated)`, fill = Race), bins = 10)
shapiro.test(race_df$`Death Rate (estimated)`)

race_df %>% 
  ggplot() +
  geom_histogram(aes(x = `Ratio Observed/ Expected`, fill = Race), bins = 10)
shapiro.test(race_df$`Ratio Observed/ Expected`)
```
Both histograms show a positively skewed distribution.

```{r}
race_df %>% 
  ggplot() +
  geom_histogram(aes(x = log(`Death Rate (estimated)` + 1), fill = Race) , bins = 10)
shapiro.test(log(race_df$`Death Rate (estimated)` + 1))

race_df %>% 
  ggplot() +
  geom_histogram(aes(x = log(`Ratio Observed/ Expected` + 1), fill = Race) , bins = 10)
shapiro.test(log(race_df$`Ratio Observed/ Expected` + 1))
```
The log transformation of Death Rate better fit, but there is still sufficient evidence to reject normality.

```{r}
race_df_DR <- race_df %>% 
  filter(`Death Rate (estimated)` > 0)
b_DR <-  boxcox(lm(race_df_DR$`Death Rate (estimated)` ~ 1))
lambda_DR <- b_DR$x[which.max(b_DR$y)]

race_df_ROE <- race_df %>% 
  filter(`Ratio Observed/ Expected` > 0)

b_ROE <-  boxcox(lm(race_df_ROE$`Ratio Observed/ Expected` ~ 1))
lambda_ROE <- b_ROE$x[which.max(b_ROE$y)]
lambda_ROE 
```

We'll create a function to carry out the box cox transformation [cite](https://stats.stackexchange.com/questions/1444/how-should-i-transform-non-negative-data-including-zeros)
```{r}
box_cox <- function(x, lambda) {

    eps <- 0.00001
    if (abs(lambda) < eps)
        log(x)
    else
        (x ^ lambda - 1) / lambda

}
```

Box Cox transformation of DR and results of shapiro test.
```{r}
race_df_DR %>%
  mutate(BOX = box_cox(`Death Rate (estimated)`, lambda_DR)) %>%
  ggplot() + geom_histogram(aes(x= BOX, fill = Race), bins = 10) + facet_wrap(. ~ Race)

race_df_DR <- race_df_DR %>%
  mutate(BOX = box_cox(`Death Rate (estimated)`, lambda_DR))

race_df_DR %>%
  pull(BOX) %>% 
  shapiro.test()
##Reject normality

race_df_DR %>%
  filter(Race == "Black") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Fail to reject normality

race_df_DR %>%
  filter(Race == "White") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Reject normality

race_df_DR %>%
  filter(Race == "Hispanic") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Fail to reject normality

race_df_DR %>%
  filter(Race == "Multiple") %>% 
  pull(BOX) %>% 
  shapiro.test()
##Fail to reject normality

par(mfrow=c(1,3))
plot(lm(race_df_DR$BOX ~ 1))
```
Successful transformation of death rate among of Black, Hispanic, and Multiple racial backgrounds persons.

Box Cox transformation of ROE and results of Shapiro test.
```{r}
race_df_ROE %>%
  mutate(BOX = box_cox(`Ratio Observed/ Expected`, lambda_ROE)) %>%
  ggplot() +
  geom_histogram(aes(x= BOX, fill = Race), bins = 10) +
  facet_wrap(. ~ Race) +
  theme_classic() +
  theme(legend.position = "none") +
  labs(x = "Box Cox Transformation of Observed/ Expected HIV Mortality Ratio", y = "Count")

ggsave(here("results/box_hist.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
race_df_ROE <- race_df_ROE %>%
  mutate(BOX = box_cox(`Ratio Observed/ Expected`, lambda_ROE))

race_df_ROE %>%
  pull(BOX) %>% 
  shapiro.test()
## Fail to reject normality

race_df_ROE %>%
  filter(Race == "Black") %>% 
  pull(BOX) %>% 
  shapiro.test()
## Fail to reject normality

race_df_ROE %>%
  filter(Race == "White") %>% 
  pull(BOX) %>% 
  shapiro.test()
## Reject normality

race_df_ROE %>%
  filter(Race == "Hispanic") %>% 
  pull(BOX) %>% 
  shapiro.test()
## Fail to reject normality

race_df_ROE %>%
  filter(Race == "Multiple") %>% 
  pull(BOX) %>% 
  shapiro.test() 
## Fail to reject normality

par(mfrow=c(1,3))
plot(lm(race_df_ROE$BOX ~ 1))
##I saved the plot above manually because I did not know how to save all three plots in one image aside from exporting.
```

The BOX transformation of ROE was better than that of DR, so let's use that one... but before wrapping up, let's go back and pull the observations that were dropped from the Box Cox transformation. 
```{r}
race_drop <- race_df %>% 
  filter(`Ratio Observed/ Expected` == 0)

race_new <- full_join(race_df_ROE, race_drop)

##We need to replace the NA with a value in order to run spatial tests later. We will replace with 0. 
race_new <-  race_new %>%
  mutate(BOX = replace(BOX, is.na(BOX),0))
```

```{r}
saveRDS(race_new, here("data/processed_data/race_new.rds"))
```