### 2023-04-08 code/analysis_code/Race/p2_PCA_race.Rmd

```{r message=FALSE, warning=FALSE}
library(ggcorrplot)
library(FactoMineR)
library(factoextra)
library(spatialreg)
library(MASS)
library(here)
library(tidyverse)

df <- readRDS(here("data/processed_data/race_new.rds"))
```

CREATING PCA FUNCTION
```{r}
pca_function <- function(race) {   
  # filter race_df by the selected race/ethnicity
  df <- df[df$Race == race,]

#Selecting variables
num_race <- df[,c(9:10, 16:25)]

#normalization
norm_race <- scale(num_race)
print(head(norm_race))

#setting up correlation matrix
corr_matrix_race <- cor(norm_race)
print(ggcorrplot(corr_matrix_race))

#applying pca
pca_race <- princomp(corr_matrix_race)
print(summary(pca_race))
#pca loadings
print(pca_race$loadings[, 1:2])

#Scree plot
print(fviz_eig(pca_race, addlabels = TRUE))
# Biplot of the attributes
print(fviz_pca_var(pca_race, col.var = "darkblue"))

#Contribution of each variable 
print(fviz_cos2(pca_race, choice = "var", axes = 1:2))

# Biplot of the attributes
print(fviz_pca_var(pca_race, col.var = "cos2",
             gradient.cols = c("darkblue", "orange", "green"),
             repel = TRUE, title = race))
# Variables in green are a good representation of the variable on that component.
}
```

```{r}
pca_function("Black")
ggsave(here("results/pca_black.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
pca_function("Hispanic")
ggsave(here("results/pca_hispanic.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
pca_function("Multiple")
ggsave(here("results/pca_multiple.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
pca_function("White")
ggsave(here("results/pca_white.png"), units="in", width=10, height=7, dpi=300)
```