### 2023-04-08 code/analysis_code/archieved/Sex/2_sex.Rmd

```{r message=FALSE, warning=FALSE}
library(here)
library(tidyverse)
library(spdep)
```

```{r}
sex_df <- readRDS(here("data/processed_data/sex_new.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))
```

```{r}
male_df <- sex_df %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State),
         Sex == "Male")

#Making empty matrix
h_male <- matrix(0, nrow(male_df), nrow(male_df))

#filling matrix with '1' for neighbors
h_male[cbind(match(us_nabs$StateCode, male_df$`State Abbreviation`), 
             match(us_nabs$NeighborStateCode,
                   male_df$`State Abbreviation`))] <- 1

nablist_male <- mat2listw(h_male, style = "W")
```

```{r}
#carrying out Moran's test - global
par(mfrow=c(1,2))
moran.test(male_df$BOX, 
           nablist_male,
           zero.policy = TRUE, 
           randomisation = FALSE)
moran.plot(male_df$BOX, 
           nablist_male, 
           zero.policy = TRUE)

#carrying out local Moran's test 
lisa_male <-localmoran(male_df$`Ratio Observed/Expected`,  nablist_male)
summary(lisa_male[,1])
```

Doing the same for females...
```{r}
female_df <- sex_df %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State),
         Sex == "Female") %>% 
  mutate(BOX = ifelse(is.na(BOX),0, BOX)) # replacing NA with zeros so that we can run the Moran's test.

h_female <- matrix(0, nrow(female_df), nrow(female_df))
h_female[cbind(match(us_nabs$StateCode, 
                     female_df$`State Abbreviation`), 
               match(us_nabs$NeighborStateCode, 
                     female_df$`State Abbreviation`))] <-1
nablist_female <- mat2listw(h_female, style ="W")

par(mfrow=c(1,2))
moran.test(female_df$BOX,
           nablist_female,
           zero.policy = TRUE, 
           randomisation = FALSE)

moran.plot(female_df$BOX, 
           nablist_female,
           zero.policy = TRUE)

lisa_female <-localmoran(female_df$`Ratio Observed/Expected`,  nablist_female)
summary(lisa_female[,1])
```