### 2023-04-08 readme.qmd

---
title: "Read Me"
format: html
editor: visual
---

This respository contains data and code to reproduce my research. Raw data is store in the `data` subfolder `raw_data`, while processed data produced by wrangling is stored under `processed_data`.

Steps to reproduce this project are:

1.  Open the R project `dataanalysis-template`
2.  Open the folder code and start by running files in `processing_code`. The files are numbered sequentially, but `1_cleaning_2020` is the only file in this folder that must be run first. The others may be run in any order.
3.  Open the folder code and start by running files in `analysis_code` , which contains sub-folders for running the analysis for race/ethnicity `race` and `sex` separately. `race` and `sex` may be run independently of each other, but contain numbers to indicate the order which these files should be run.
4.  Data files produced by these files are store in the `data` subfolder `processed_data`