### 2023-04-21 code/analysis_code/Race/p2_PCA_race.Rmd

```{r message=FALSE, warning=FALSE}
library(ggcorrplot)
library(FactoMineR)
library(factoextra)
library(here)
library(knitr)
library(tidyverse)
library(broom)

df <- readRDS(here("data/processed_data/race_new.rds"))
```

PCA
```{r}
df <- df %>%
    select_if(~ !any(is.na(.)))

  #separating the y variables
df_response <- df %>%
    select(contains(c("State", "Race", "Death", "Population", "BOX", "YJ", "Ratio", "outcome_DR")))

  # separating x variables and scaling
df_predictor <- df %>% 
    select(contains(c("State", "Race", "Percent", "Gini", "Median", "Broadband"))) %>% 
    #select(!contains(c("Virally", "Received", "Late"))) %>% 
    mutate(across(where(is.numeric), scale))

#Setting up correlation matrix
corr_matrix <- df_predictor %>% 
    select(where(is.numeric)) %>% 
    cor()

  #setting up correlation matrix 
print(ggcorrplot(corr_matrix))
ggsave(here("results/corr_mat.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
#applying pca
pca <- princomp(corr_matrix)
print(summary(pca))

#pca loadings of principal components 1 and 2
print(pca$loadings[, 1:2])

#Scree plot
print(fviz_eig(pca, addlabels = TRUE))
ggsave(here("results/scree_plot.png"), units="in", width=10, height=7, dpi=300)
```
Since we originally had 11 variables, we should only look at components that explain more than 9.1% of variation. We can see from the scree plot that principal component 1 explained 63.8% of variation, and components 1 and 2 explained 84.7% of the variation, and components 1 - 3 explained 94.8%. Component 3 just  barely passed the threshold with 10.1%. All other principal components explains less that 6.2% of variation, so they will excluded.

```{r}
#Contribution of each variable 
print(fviz_cos2(pca, choice = "var", axes = 1:2))

# Biplot of the attributes
print(fviz_pca_var(pca, col.var = "cos2",
                   gradient.cols = c("darkblue", "orange", "green"),
                   repel = TRUE, title = "PCA"
                   ))
# Variables in green are a good representation of the variable on that component.

#Creating object to store scaled dataframes
df_scaled <- full_join(df_response, df_predictor, by = join_by("State", "State Abbreviation", "States with HIV-Specific Criminal Laws/Statutes", "States That Have Prosecuted PLWH", "Race"))
```

```{r}
saveRDS(df_scaled, here("data/processed_data/df_scaled.rds"))
```