### 2023-04-21 outputs_code.Rmd

#Table
library(kableExtra)
```{r}
df <- readRDS(here("data/processed_data/race_new.rds"))

df %>% 
  group_by(Race) %>% 
  summarize(tot_deaths = sum(`Death Cases`),
            tot_population = sum(`Population`),
            overall_rate = round(tot_deaths/ tot_population *1e5,1)) %>% 
  ungroup() %>% 
  mutate(percent_by_race = round(tot_deaths/sum(tot_deaths)*100, 1),
         ratio = round(overall_rate/overall_rate[Race =="White"],1)) %>%
  kable(col.names = c("Race/ Ethnicity",
                      "HIV-Related Deaths (n)",
                      "Population \n (>12 years old)",
                      "Death Rate \n (per 100,000)",
                      "Percent",
                      "Ratio"),
        caption = "HIV-Related Deaths Summary by Race/ Ethncity, United States, 2020",
        align = "c") %>%
  kable_minimal() %>%
  kable_styling(font_size = 50) %>% 
  save_kable(here("results/test.png"))
# using white persons as reference group
```