### 2023-04-25 code/analysis_code/1_EDA_alt.Rmd
modified outcome variable to MRR

```{r message=FALSE, warning=FALSE}
library(here)
library(spdep)
library(GGally)
library(skimr)
library(ggfortify)
library(MASS)
library(ggcorrplot)
library("FactoMineR")
library(tidyverse)
library(broom)
library(knitr)
library(ggpubr)
```

https://github.com/ubikuity/List-of-neighboring-states-for-each-US-state
List-of-neighboring-states-for-each-US-state

```{r}
race_sub <- readRDS(here("data/processed_data/race_df_alt.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subsetting to only contiguous states.
race_sub <- race_sub %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))

us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))
```

Let's create a variable for mortality rate ratio of Black/ White
```{r}
race_sub <- race_sub %>%
  group_by(State) %>%
  mutate(MRR = `Death Rate (estimated)`[Race == "Black"] / `Death Rate (estimated)`[Race == "White"])
```

Let's start by looking at the overall HIV-related death rate by race in the United States.
```{r}
race_sub %>% 
  group_by(Race) %>% 
  summarize(tot_deaths = sum(`Death Cases`),
            tot_population = sum(`Population`),
            overall_rate = round(tot_deaths/ tot_population *1e5,1)) %>% 
  ungroup() %>% 
  mutate(percent_by_race = round(tot_deaths/sum(tot_deaths)*100, 1),
         ratio = round(overall_rate/overall_rate[Race =="White"],1)) %>%
  kable(col.names = c("Race/ Ethnicity",
                      "HIV-Related Deaths (n)",
                      "Population (>12 years old)",
                      "Death Rate (per 100,000)",
                      "Percent",
                      "Ratio"))
# using white persons as reference group
```

We can seen that Black persons accounts for the 43.5% of HIV-related deaths in United States, and the rate of HIV-related deaths among this population was 7.2 times high than the rate of HIV-related deaths among White persons.

Next, let's take a look at our response variable
Univariate Analysis

Let's take a look at some box plots of death rate by races and state. 
We'll start by looking at death rate given by the AIDSVu dataset. 
```{r}
race_sub %>% 
  ggplot() +
  geom_boxplot(aes(y = `Death Rate (estimated)`, x = Race, group = Race)) + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio, United States, 2020", y = "Ratio", x = "Race/ Ethnicity") + theme(legend.position = "none") 
```

Histogram
```{r}
race_sub %>% 
  filter(Race == "Black") %>% 
  ggplot() +
  geom_histogram(aes(x = `MRR`), bins = 10) + 
  theme(legend.position = "none") +
  labs(title = "Histogram of Black-to-White HIV-Related Mortality Rate Ratio, United States, 2020", y = "n", x = "Ratio") + theme_classic()
ggsave(here("results/MRR_hist.png"), units="in", width=10, height=7, dpi=300)

race_sub %>% 
   filter(Race == "Black") %>% 
  pull(`MRR`) %>% 
  shapiro.test()

race_sub %>% 
  ggplot() +
  geom_boxplot(aes(y = `Death Rate (estimated)`, x = Race, group = Race)) + 
  theme(legend.position = "none") +
  labs(title = "HIV-Related Death Rate, by Race, United States, 2020", y = "Rate (per 100,000)", x = "Race") + theme_classic()
ggsave(here("results/dr_race_bp.png"), units="in", width=10, height=5, dpi=300)

race_sub %>% 
  ggplot() +
  geom_bar(aes(y = `Death Rate (estimated)`, x = `State Abbreviation`, group = `State Abbreviation`, fill = Race),
           stat = "identity", position = "dodge") + theme_classic() + 
  theme(legend.position = "bottom") +
  labs(title = "HIV-Related Death Rate of Black and White Persons, by State, United States, 2020", y = "Rate (per 100,000)", x = "State")
ggsave(here("results/dr_state.png"), units="in", width=12, height=5, dpi=300)

```

Let's move on to
Bivariate Analysis
```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Living in Poverty`)) +
  geom_smooth(method = "lm") + 
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Living in Poverty") + 
  stat_cor(method = "pearson") 
#  pass
ggsave(here("results/bi_pov.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Unemployed`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Unemployed") + 
  stat_cor(method = "pearson") 
ggsave(here("results/bi_unemployed.png"), units="in", width=10, height=7, dpi=300)
#  did not pass 
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Gini Coefficient`)) +
    geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Gini Coefficient") +
  stat_cor(method = "pearson") 
ggsave(here("results/bi_gini.png"), units="in", width=10, height=7, dpi=300)
# did not "pass"
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Less than High School Education`)) +
  geom_smooth(method = "lm") + 
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Less than High School Education") +
  stat_cor(method="pearson")
#  did not pass
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Living with Severe Housing Cost Burden`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Living with Severe Housing Cost Burden") + 
  stat_cor(method="pearson")
#  did not pass
ggsave(here("results/bi_housing.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Without an Internet subscription`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Without an Internet Subscription") + 
  stat_cor(method="pearson")
# "pass"
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Uninsured Under 65 Years Old`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Uninsured Under 65 Years Old") + 
  stat_cor(method="pearson")
# "pass"
ggsave(here("results/bi_uninsured.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Living with Food Insecurity`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Living with Food Insecurity") + 
  stat_cor(method="pearson")
# pass"
ggsave(here("results/bi_foodinsecure.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Median Household Income`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Median Household Income") + 
  stat_cor(method="pearson")
# "pass"
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percentage Diagnosed Late`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percentage Diagnosed Late") + 
  stat_cor(method="pearson")
# "pass"
ggsave(here("results/bi_latedx.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percentage Virally Suppressed`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percentage Virally Suppressed") + 
  stat_cor(method="pearson")
#did not pass
```
Now that we have an idea of the relationship between our outcome and potential predictors, next I need to account for the fact that these variables a likely have some degree of multicollinearity. In the next step we will be using PCA to scale our predictors.

```{r}
pos_predictors <- race_sub %>% 
  select(c("State", 
           "Race",
           "Percent Living in Poverty",
           "Percent Without an Internet subscription",
           "Percent Uninsured Under 65 Years Old",
           "Percent Living with Food Insecurity",
           "Percent Living with Severe Housing Cost Burden",
           "Median Household Income",
           "Percentage Diagnosed Late"))
```

```{r}
saveRDS(pos_predictors, here("data/processed_data/pos_predictors.rds"))
saveRDS(race_sub, here("data/processed_data/race_alt.rds"))
```