### 2023-04-25 code/analysis_code/archieved/Race/p3_model_race.Rmd

```{r message=FALSE, warning=FALSE}
library(here)
library(MASS)
library(spatialreg)
library(spdep)
library(tidyverse)
```

```{r}
df <- readRDS(here("data/processed_data/df_scaled.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subset to contiguous states
us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))

df <- df %>% 
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))
```

```{r}
states <- data.frame(state.name)
states$region <- case_when(
  states$state.name %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania") ~ "Northeast",
  states$state.name %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin") ~ "Midwest",
  states$state.name %in% c("Alabama", "Arkansas", "Delaware", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia") ~ "South",
  states$state.name %in% c("Arizona", "California", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming") ~ "West"
)

df <- full_join(df, states, by = join_by(State == state.name))
```

subset by race/ethnicity
```{r}
black_df <- df %>% 
  filter(Race == "Black")

hispanic_df <- df %>% 
  filter(Race == "Hispanic")

white_df <- df %>% 
  filter(Race == "White")
```

Setting up functions for models
```{r}
 #multiple linear regression
mult_function <- function(df) {
  mult_race <- glm(outcome_DR ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`, #economic stability,
                     data = df,
                     family = "gaussian")
  print(summary(mult_race))
    ##pull fitted values
df$mult_fit <- mult_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
mult_aic <- function(df) {
  mult_race <- glm(outcome_DR ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`, #economic stability,
                     data = df,
                     family = "gaussian")
aic <- AIC(mult_race)

return(aic)
}
```

```{r}
  #poisson
pois_function <- function(df) {
  pois_race <- glm(`Death Cases`~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                    offset(log(`Population`)),
                     data = df,
                     family = "poisson")
  print(summary(pois_race))

      ##pull fitted values
df$pois_fit <- pois_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
pois_aic <- function(df) {
  pois_race <- glm(`Death Cases`~ 
                         `Gini Coefficient` + # estimate income inequality
                         `Percent Unemployed` +
                         `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                    offset(log(`Population`)),
                     data = df,
                     family = "poisson")
aic <- AIC(pois_race)
return(aic)
}
```

```{r}
  #Negative binomial
nb_function  <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ 
                        `Gini Coefficient` + # estimate income inequality
                         `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                     offset(log(`Population`)),
                            data = df)
  print(summary(neg_bin_race))

      ##pull fitted values
df$nb_fit <- neg_bin_race$fitted.values

return(df)

}

## function to fit GLM and pull AIC
nb_aic <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ 
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden` + #economic stability
                     offset(log(`Population`)),
                            data = df)
aic <- AIC(neg_bin_race)

return(aic)

}  
```

```{r}
#Fitting CAR model (requires neighbors)
car_function <- function(df) {   
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

  car_race <- spautolm(outcome_DR ~                      
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`,#economic stability,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
  print(summary(car_race))

  ##pull fitted values
df$car_fit <- car_race$fit$fitted.values

return(df)

}

car_aic <- function(df) {
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

  car_race <- spautolm(outcome_DR ~                      
                        `Gini Coefficient` + # estimate income inequality
                        `Percent Unemployed` +
                        `Percent Less than High School Education` +  #estimate education access
                        `Percent HIV Prevalence` +
                        `Percent Living with Severe Housing Cost Burden`,#economic stability,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
aic <- AIC(car_race)
return(aic)

} 
```

Making a function to pull AIC and organize into new dataframe
```{r}
aic_df <- function(x, mod_type, race) {
  x <- tibble(x)
  x <- x %>% 
    mutate(Race = race, 
           Model = mod_type,
           AIC = round(x,2)) %>% 
    select(Race, Model, AIC)
}
```

Fitting multiple linear regression models
```{r}
black_df <- mult_function(black_df)

black_mult <- mult_aic(black_df)
black_mult <- aic_df(black_mult, "GLM", "Black")
```

```{r}
hispanic_df <- mult_function(hispanic_df)

hispanic_mult <- mult_aic(hispanic_df)
hispanic_mult <- aic_df(hispanic_mult, "GLM", "Hispanic")
```

```{r}
white_df <- mult_function(white_df)
## SINCE THE OUTCOME IN NOT NORMALLY DISTRIBUTED, WE CANNOT USE GLM 
white_df <- white_df %>% mutate(mult_fit = NA)

white_mult <- mult_aic(white_df)
white_mult <- aic_df(white_mult, "GLM", "White")

white_mult <- white_mult %>% mutate(AIC = NA)
```

```{r}
mult_mod <- list(black_mult, hispanic_mult, white_mult)

mult_mod <- reduce(mult_mod, full_join)
```

Fitting Poisson models
```{r}
black_df <- pois_function(black_df)

black_pois <- pois_aic(black_df)
black_pois <- aic_df(black_pois, "Poisson", "Black")
```

```{r}
hispanic_df <- pois_function(hispanic_df)

hispanic_pois <- pois_aic(hispanic_df)
hispanic_pois <- aic_df(hispanic_pois, "Poisson", "Hispanic")
```

```{r}
white_df <- pois_function(white_df)

white_pois <- pois_aic(white_df)
white_pois <- aic_df(white_pois, "Poisson", "White")
```

Storing Poisson AIC
```{r}
pois_mod <- list(black_pois, hispanic_pois, white_pois)

pois_mod <- reduce(pois_mod, full_join)
```

Fitting Negative binomial models
```{r}
black_df <- nb_function(black_df)

black_nb <- nb_aic(black_df)
black_nb <- aic_df(black_nb, "Negative Binomial", "Black")
```

```{r}
hispanic_df <- nb_function(hispanic_df)

hispanic_nb <- nb_aic(hispanic_df)
hispanic_nb <- aic_df(hispanic_nb, "Negative Binomial", "Hispanic")
```

```{r}
white_df <- nb_function(white_df)

white_nb <- nb_aic(white_df)
white_nb <- aic_df(white_nb, "Negative Binomial", "White")
```

```{r}
nb_mod <- list(black_nb, hispanic_nb, white_nb)

nb_mod <- reduce(nb_mod, full_join)
```
The NB model accounting for overdispersion fits better than the poisson model.

Fitting CAR models
```{r}
black_df <- car_function(black_df)

black_car <- car_aic(black_df)
black_car <- aic_df(black_car, "CAR", "Black")
```

```{r}
hispanic_df <- car_function(hispanic_df)

hispanic_car <- car_aic(hispanic_df)
hispanic_car <- aic_df(hispanic_car, "CAR", "Hispanic")
```

```{r}
white_df <- car_function(white_df)
## SINCE THE OUTCOME IN NOT NORMALLY DISTRIBUTED, WE CANNOT USE CAR
white_df <- white_df %>% mutate(car_fit = NA)

white_car <- car_aic(white_df)
white_car <- aic_df(white_car, "CAR", "White")
white_car <- white_car %>% mutate(AIC = NA)
```

```{r}
car_mod <- list(black_car, hispanic_car, white_car)

car_mod <- reduce(car_mod, full_join)
```

```{r}
mod_aic <- list(mult_mod, pois_mod, nb_mod, car_mod)

mod_aic <- reduce(mod_aic, full_join)
```

Join the data frames with all the fits.
```{r}
df <- list(black_df, hispanic_df, white_df)

df <- reduce(df, full_join)

saveRDS(df, here("data/processed_data/df_fit.rds"))
```
save the AIC dataframe
```{r}
saveRDS(mod_aic, here("data/processed_data/aic.rds"))
```
The model accounting for spatial variation provide the best fit.

Creating function that subsets by specified race, and compute/ plots Moran's I and lisa.
```{r}
spacy_function <- function(df) {

  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

 # plot Moran's I
  print(moran.test(df$outcome_DR, nablist_race_df, zero.policy = TRUE, randomisation = FALSE))
  moran.plot(df$outcome_DR, nablist_race_df, zero.policy = TRUE)

  # calculate and print LISA statistics
  lisa_race <- localmoran(df$outcome_DR, nablist_race_df)
  summary(lisa_race[,1])
}
```

```{r}
spacy_function(black_df)
```

```{r}
png(here("results/moran_black.png"))
sink(here("results/spacy_summary_black.txt"))
spacy_function(black_df)
sink()
dev.off()
```

```{r}
spacy_function(hispanic_df)
```

```{r}
png(here("results/moran_hispanic.png"))
sink(here("results/spacy_summary_hispanic.txt"))
spacy_function(hispanic_df)
sink()
dev.off()
```

```{r}
save.image(here("data/processed_data/race_env.rds"))
```