### 2023-04-25 code/analysis_code/3_model_alt.Rmd

```{r message=FALSE, warning=FALSE}
library(here)
library(MASS)
library(spatialreg)
library(spdep)
library(tidyverse)
```

```{r}
df <- readRDS(here("data/processed_data/df_scaled_alt.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subset to contiguous states
us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))

```

```{r}
states <- data.frame(state.name)
states$region <- case_when(
  states$state.name %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania") ~ "Northeast",
  states$state.name %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin") ~ "Midwest",
  states$state.name %in% c("Alabama", "Arkansas", "Delaware", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia") ~ "South",
  states$state.name %in% c("Arizona", "California", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming") ~ "West"
)

df <- full_join(df, states, by = join_by(State == state.name))

df <- df %>% 
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))
```

Setting up functions for models
```{r}
 #multiple linear regression
mult_function <- function(df) {
  mult_race <- glm(MRR ~ 
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old`,
                      #  `Percent Living with Food Insecurity`,
                     data = df,
                     family = "gaussian")
  print(summary(mult_race))
    ##pull fitted values
df$mult_fit <- mult_race$fitted.values
return(df)
}

## function to fit GLM and pull AIC
mult_aic <- function(df) {
  mult_race <- glm(MRR ~ 
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old`,
                      #  `Percent Living with Food Insecurity`,
                     data = df,
                     family = "gaussian")
aic <- AIC(mult_race)

return(aic)
}
```

```{r}
df <- mult_function(df)

df_mult <- mult_aic(df)
```

```{r}
  #poisson
pois_function <- function(df) {

pois_race <- glm(`Death Cases` ~ Race + region +
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                      #  `Percent Living with Food Insecurity`  +
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))
  print(summary(pois_race))

      ##pull fitted values
df$pois_fit <- pois_race$fitted.values

return(df)

}

## function to fit GLM and pull AIC
pois_aic <- function(df) {
  pois_race <- glm(`Death Cases` ~  Race + region +
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                      #  `Percent Living with Food Insecurity`  +
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))
aic <- AIC(pois_race)
return(aic)
}
```

```{r}
df <- pois_function(df)

df_pois <- pois_aic(df)
```

```{r}
  #Negative binomial
nb_function  <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ Race + region +
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                      #  `Percent Living with Food Insecurity`  +
                     offset(log(`Population`)),
                            data = df)
  print(summary(neg_bin_race))

      ##pull fitted values
df$nb_fit <- neg_bin_race$fitted.values

return(df)

}

## function to fit GLM and pull AIC
nb_aic <- function(df) {
  neg_bin_race <- glm.nb(`Death Cases` ~ Race + region +
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old` +
                      #  `Percent Living with Food Insecurity`  +
                     offset(log(`Population`)),
                            data = df)
aic <- AIC(neg_bin_race)

return(aic)

}  
```

```{r}
df <- nb_function(df)

df_nb <- nb_aic(df)
```

```{r}
#Fitting CAR model (requires neighbors)
car_function <- function(df) {   
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

  car_race <- spautolm(MRR ~                      
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old`,
                      #  `Percent Living with Food Insecurity`,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
  print(summary(car_race))

  ##pull fitted values
df$car_fit <- car_race$fit$fitted.values

return(df)

}

car_aic <- function(df) {
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

  car_race <- spautolm(MRR ~                      
                       # `Percent Living in Poverty` +
                        `Gini Coefficient` + 
                        `Percent Without an Internet subscription` +
                        `Percent Uninsured Under 65 Years Old`,
                      #  `Percent Living with Food Insecurity`,
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")
aic <- AIC(car_race)
return(aic)

} 
```

```{r}
df <- car_function(df)

df_car <- car_aic(df)
```

Organizing mod_AIC into new dataframe
```{r}
aic_df <- data.frame(AIC = c(df_mult, df_pois, df_nb, df_car),
                     Model = c("Multiple Linear Regression", "Poisson", "Negative Binomial", "Conditional Autoregressive"),
                     Outcome = c("Ratio", "Count", "Count", "Ratio"))
aic_df <- aic_df %>% mutate(AIC  = round(AIC, 2))

```

```{r}
library(kableExtra)
library(knitr)
table_fit <- aic_df %>%
  select(c(Model, Outcome, AIC)) %>% 
  kable(caption = "<center>Model Fit</center>",
        align = "c") %>%
  kable_minimal() %>%
  kable_styling(font_size = 30) %>%
gsub("font-size: initial !important;",
"font-size: 30pt !important;",
.)  %>% 
  landscape()

table_fit %>%  
  save_kable(here("results/table_fit.png"))

```

```{r}
saveRDS(df, here("data/processed_data/df_fit_alt.rds"))
```

Creating function that subsets by specified race, and compute/ plots Moran's I and lisa.
```{r}
spacy_function <- function(df) {

  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

 # plot Moran's I
  print(moran.test(df$MRR, nablist_race_df, zero.policy = TRUE, randomisation = FALSE))
  moran.plot(df$MRR, nablist_race_df, zero.policy = TRUE)

  # calculate and print LISA statistics
  lisa_race <- localmoran(df$MRR, nablist_race_df)
  summary(lisa_race[,1])
}
```

```{r}
spacy_function(df)
```

```{r}
png(here("results/moran_mrr.png"))
spacy_function(df)
dev.off()
```