### 2023-04-30 code/analysis_code/2_pca_alt.Rmd
hopefully this works...

```{r message=FALSE, warning=FALSE}
library(ggcorrplot)
library(FactoMineR)
library(factoextra)
library(here)
library(knitr)
library(tidyverse)
library(broom)

race_sub <- readRDS(here("data/processed_data/race_alt.rds"))
pos_predictors <- readRDS(here("data/processed_data/pos_predictors.rds"))

```

PCA
```{r}
race_sub <- race_sub %>%
    select_if(~ !any(is.na(.))) %>% 
  #filter(Race == "Black") %>% 
  ungroup()

  #separating the y variables
df_response <- race_sub %>%
    select(contains(c("State","Region", "Ratio", "MRR", "Race", "Death", "Population")))

  # pulling variables that were correlated with MRR, and then scaling
pos_predictors_scaled <- pos_predictors %>% 
    ungroup() %>% 
    mutate(across(where(is.numeric), scale))

#Setting up correlation matrix
corr_matrix <- pos_predictors_scaled %>% 
    select(c(3:7)) %>% 
    cor()

  #setting up correlation matrix 
pca_1 <-print(ggcorrplot(corr_matrix))
#ggsave(here("results/corr_mat.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
#applying pca
pca <- princomp(corr_matrix)
print(summary(pca))

#pca loadings of principal components 1 and 2
print(pca$loadings[, 1:3])

#Scree plot
pca_2 <-print(fviz_eig(pca, addlabels = TRUE))
#ggsave(here("results/scree_plot.png"), units="in", width=10, height=7, dpi=300)
```
Since we originally had 5 variables, we should only look at components that explain more than 20% of variation. This is includes components 1 through 3. All other principal components explains less that 8.9% of variation, so they will excluded.

```{r}
#Contribution of each variable 
pca_3 <-print(fviz_cos2(pca, choice = "var", axes = 1:2))
#ggsave(here("results/contribution_plot.png"), units="in", width=10, height=7, dpi=300)

# Biplot of the attributes
pca_4 <- print(fviz_pca_var(pca, col.var = "cos2",
                   gradient.cols = c("darkblue", "orange", "green"),
                   repel = TRUE, title = "PCA"
                   ))
#ggsave(here("results/biplot.png"), units="in", width=10, height=7, dpi=300)
# Variables in green are a good representation of the variable on that component.

#Creating object to store scaled dataframes
df_scaled <- full_join(df_response, pos_predictors_scaled, by = join_by("State", "Race"))
```

```{r}
library(gridExtra)
grid_pca <- grid.arrange(pca_1, pca_2, pca_3, pca_4, ncol = 2, nrow = 2)

ggsave(here("results/grid_pca.png"), grid_pca, units="in", width=10, height=10, dpi=300)
```

```{r}
saveRDS(df_scaled, here("data/processed_data/df_scaled_alt.rds"))
```