### 2023-05-05 code/analysis_code/1_EDA.Rmd
submitting final changes

```{r message=FALSE, warning=FALSE}
library(here)
library(ggfortify)
library(tidyverse)
library(broom)
library(knitr)
library(kableExtra)
library(ggpubr)
```

https://github.com/ubikuity/List-of-neighboring-states-for-each-US-state
List-of-neighboring-states-for-each-US-state

```{r}
race_sub <- readRDS(here("data/processed_data/race_df_alt.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subsetting to only contiguous states.
race_sub <- race_sub %>%
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))

us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))
```

Creating variable for region
```{r}
states <- data.frame(state.name)
states$Region <- case_when(
  states$state.name %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania") ~ "Northeast",
  states$state.name %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin") ~ "Midwest",
  states$state.name %in% c("Alabama", "Arkansas", "Delaware", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia") ~ "South",
  states$state.name %in% c("Arizona", "California", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming") ~ "West"
)

race_sub <- full_join(race_sub, states, by = join_by(State == state.name))

race_sub <- race_sub %>% 
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))
```

Let's start by looking at the overall HIV-related death rate by race in the United States.
```{r}
table1 <- race_sub %>% 
  group_by(Race) %>% 
  summarize(tot_deaths = sum(`Death Cases`),
            tot_population = sum(`Population`),
            overall_rate = round(tot_deaths/ tot_population *1e5,1)) %>% 
  ungroup() %>% 
  mutate(ratio = round(overall_rate/overall_rate[Race =="White"],1)) %>%
  kable(col.names = c("Race/ Ethnicity",
                      "HIV-Related Deaths (n)",
                      "Population \n (>12 years old)",
                      "Death Rate \n (per 100,000)",
                      "Mortality Rate Ratio"),
        caption = "<center>HIV-Related Deaths by Race/ Ethnicity, United States, 2020</center>",
        align = "c", format.args = list(big.mark = ",")) %>%
  kable_minimal() %>%
  kable_styling(font_size = 50)%>%
gsub("font-size: initial !important;",
"font-size: 50pt !important;",
.)  %>% 
  landscape()

table1

table1 %>%  
  save_kable(here("results/table1.png"))
# using white persons as reference group
```

```{r}
region_table <- race_sub %>% 
  group_by(Region, Race) %>% 
  summarize(tot_deaths = sum(`Death Cases`),
            tot_population = sum(`Population`),
            overall_rate = round(tot_deaths/ tot_population *1e5,1)) %>% 
  ungroup() %>% 
  group_by(Region) %>% 
  mutate(ratio = round(overall_rate/overall_rate[Race =="White"],1)) %>%
  kable(col.names = c("Region",
                      "Race/ Ethnicity",
                      "HIV-Related Deaths (n)",
                      "Population \n (>12 years old)",
                      "Death Rate \n (per 100,000)",
                      "Mortality Rate Ratio"),
        caption = "<center>HIV-Related Deaths by Region, United States, 2020</center>", format.args = list(big.mark = ",")) %>% 
  kable_minimal() %>%
  kable_styling(font_size = 50) %>%
gsub("font-size: initial !important;",
"font-size: 50pt !important;",.)  %>% 
  landscape()

region_table

region_table %>%  
  save_kable(here("results/region_table.png"))
# using white persons as reference group
```

We can seen that Black persons accounts for the 43.5% of HIV-related deaths in United States, and the rate of HIV-related deaths among this population was 7.2 times high than the rate of HIV-related deaths among White persons.

Next we'll look into the distribution of our response variable
```{r}
race_sub %>% 
  filter(Race == "Black") %>% 
  ggplot() +
  geom_histogram(aes(x = `MRR`), bins = 10) + 
  theme(legend.position = "none") +
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio, United States, 2020", y = "n", x = "Ratio") + theme_classic()
ggsave(here("results/MRR_hist.png"), units="in", width=10, height=5, dpi=300)

race_sub %>% 
   filter(Race == "Black") %>% 
  pull(`MRR`) %>% 
  shapiro.test()

```

Let's also look at the distribution of HIV-related death rate
```{r}
race_sub %>% 
  ggplot() +
  geom_histogram(aes(x = `Death Rate (estimated)`, fill = Race), bins = 20) + 
  theme(legend.position = "none") +
  labs(title = "HIV-Related Death Rate, United States, 2020", y = "n", x = "Deaths per 100,000") +
  theme_classic()
ggsave(here("results/DR_hist.png"), units="in", width=10, height=5, dpi=300)

```

Let's take a look at MRR by state and region
```{r}
race_sub %>% 
  ggplot() +
  geom_bar(aes(y = `MRR`, x = `State Abbreviation`, group = `State Abbreviation`),
           stat = "identity", position = "dodge") +
  theme_classic() + 
  theme(legend.position = "bottom") +
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio, by State, United States, 2020", y = "Ratio", x = "State")
ggsave(here("results/mrr_state.png"), units="in", width=12, height=5, dpi=300)

race_sub %>% 
  ggplot() +
  geom_boxplot(aes(y = `MRR`, x = `Region`, group = `Region`)) +
  theme_classic() + 
  theme(legend.position = "bottom") +
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio, by Region, United States, 2020", y = "Ratio", x = "Region")
ggsave(here("results/mrr_region.png"), units="in", width=12, height=5, dpi=300)
```

Let's move on to
Bivariate Analysis 
We will deem variables with a correlation coefficient of |0.15| as important, and will include these in PCA
```{r}
bi_1 <- ggplot(race_sub, aes(y = MRR, x = `Percent Living in Poverty`)) +
  geom_smooth(method = "lm") + 
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs. Percent Living in Poverty") + 
  stat_cor(method = "pearson") 
#  pass 
#ggsave(here("results/bi_pov.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Unemployed`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs. Percent Unemployed") + 
  stat_cor(method = "pearson") 
#  did not pass 
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Gini Coefficient`)) +
    geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs. Gini Coefficient") +
  stat_cor(method = "pearson") 
# did not "pass"
#ggsave(here("results/bi_gini.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Less than High School Education`)) +
  geom_smooth(method = "lm") + 
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Less than High School Education") +
  stat_cor(method="pearson")
#  did not pass
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percent Living with Severe Housing Cost Burden`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) vs.
       Percent Living with Severe Housing Cost Burden") + 
  stat_cor(method="pearson")
#  did not pass
#ggsave(here("results/bi_housing.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
bi_3 <- ggplot(race_sub, aes(y = MRR, x = `Percent Without an Internet subscription`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs.  Percent Without an Internet Subscription") + 
  stat_cor(method="pearson")
# "pass"
#ggsave(here("results/bi_no_internet.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
bi_4 <- ggplot(race_sub, aes(y = MRR, x = `Percent Uninsured Under 65 Years Old`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) 
  vs.  Percent Uninsured Under 65 Years Old") + 
  stat_cor(method="pearson")
# "pass"
#ggsave(here("results/bi_uninsured.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
bi_5 <- ggplot(race_sub, aes(y = MRR, x = `Percent Living with Food Insecurity`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs. Percent Living with Food Insecurity") + 
  stat_cor(method="pearson")
# pass"
#ggsave(here("results/bi_foodinsecure.png"), units="in", width=10, height=7, dpi=300)
```

```{r}
bi_2 <- ggplot(race_sub, aes(y = MRR, x = `Median Household Income`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs. Median Household Income") + 
  stat_cor(method="pearson")
# did not "pass"
```

```{r}
bi_6 <-ggplot(race_sub, aes(y = MRR, x = `Percentage Diagnosed Late`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR)
  vs.  Percentage Diagnosed Late") + 
  stat_cor(method="pearson")
# "pass"
```

```{r}
ggplot(race_sub, aes(y = MRR, x = `Percentage Virally Suppressed`)) +
  geom_smooth(method = "lm") +
  geom_point() + 
  labs(title = "Black-to-White HIV-Related Mortality Rate Ratio (MRR) 
  vs.    Percentage Virally Suppressed") + 
  stat_cor(method="pearson")
#did not pass
```

```{r}
library(gridExtra)
bi_g <- grid.arrange(bi_1, bi_2, bi_4, bi_5, bi_3, bi_6, ncol = 2, nrow = 3)

ggsave(here("results/grid_bivar.png"), bi_g, units="in", width=10, height=10, dpi=300) 
```

Now that we have an idea of the relationship between our outcome and potential predictors, next I need to account for the fact that these variables a likely have some degree of multicollinearity. In the next step we will be using PCA to scale our predictors.

```{r}
pos_predictors <- race_sub %>%
  dplyr::select(c("Median Household Income",
           "Percent Living in Poverty",
           "Percent Without an Internet subscription",
           "Percent Uninsured Under 65 Years Old",
           "Percent Living with Food Insecurity",
           "Percentage Diagnosed Late")) %>% 
  names()
```

```{r}
saveRDS(pos_predictors, here("data/processed_data/pos_predictors.rds"))
saveRDS(race_sub, here("data/processed_data/race_alt.rds"))
```