### 2023-05-05 code/analysis_code/3_model_fitting.Rmd

```{r message=FALSE, warning=FALSE}
library(here)
library(mfx)
library(spatialreg)
library(spdep)
library(tidyverse)
library(AER)
```

```{r}
df <- readRDS(here("data/processed_data/race_alt.rds"))
prin_comp1 <- readRDS(here("data/processed_data/prin_comp1.rds"))
prin_comp2 <- readRDS(here("data/processed_data/prin_comp2.rds"))
us_nabs <- readRDS(here("data/processed_data/us_nabs.rds"))

#Subset to contiguous states
us_nabs <- us_nabs %>%
  filter(!grepl("AK", StateCode),
         !grepl("HI", StateCode), 
         !grepl("DC", StateCode))

```

```{r}
states <- data.frame(state.name)
states$region <- case_when(
  states$state.name %in% c("Connecticut", "Maine", "Massachusetts", "New Hampshire", "Rhode Island", "Vermont", "New Jersey", "New York", "Pennsylvania") ~ "Northeast",
  states$state.name %in% c("Illinois", "Indiana", "Iowa", "Kansas", "Michigan", "Minnesota", "Missouri", "Nebraska", "North Dakota", "Ohio", "South Dakota", "Wisconsin") ~ "Midwest",
  states$state.name %in% c("Alabama", "Arkansas", "Delaware", "Florida", "Georgia", "Kentucky", "Louisiana", "Maryland", "Mississippi", "North Carolina", "Oklahoma", "South Carolina", "Tennessee", "Texas", "Virginia", "West Virginia") ~ "South",
  states$state.name %in% c("Arizona", "California", "Colorado", "Idaho", "Montana", "Nevada", "New Mexico", "Oregon", "Utah", "Washington", "Wyoming") ~ "West"
)

df <- full_join(df, states, by = join_by(State == state.name))

df <- df %>% 
  filter(!grepl("Alaska", State),
         !grepl("Hawaii", State))

# reorder race variable to make White the reference group for NB
df <- df %>% 
  mutate(Race = factor(Race, levels = c("White","Black")))
```

# multiple linear regression
```{r}
mult_race <- glm(MRR ~ 
                        prin_comp1 + prin_comp2,
                     data = df,
                     family = "gaussian")
mult_sum <-   summary(mult_race)
mult_sum

#storing values in table and computing confidence interval
mult_table <- as.tibble(coef(mult_sum)) %>% 
  mutate(low = round(`Estimate` - 1.96 * `Std. Error`, 2),
         high = round(`Estimate` + 1.96 * `Std. Error`, 2),
        value = c("Intercept",
                   "PC1",
                   "PC2")) %>% 
  relocate(value, .before = Estimate) %>%
  relocate(c(low, high), .before = `t value`) %>% 
  unite("95% CI", c("low", "high"), sep = ", ", na.rm = TRUE)

  ##pulling and storing fitted values
df$mult_fit <- mult_race$fitted.values

## pulling AIC
aic_mult <- AIC(mult_race)
```

#poisson
```{r}
pois_race <- glm(`Death Cases` ~ Race + region +
                        prin_comp1 + prin_comp2 + 
                     offset(log(`Population`)),
                     data = df,
                     family = poisson(link = "log"))

#Testing for overdispersion
dispersiontest(pois_race, trafo=1)
#Evidence of overdisperion with c = 17.35. Will instead utilize negative binomial regression, which can account for overdispersion.
```
  #Negative binomial that outputs rate ratios
```{r}
neg_bin_race <- negbinirr(`Death Cases` ~  Race +
                     prin_comp1  + prin_comp2 +
                      offset(log(`Population`)),
                             data = df)
nb_sum <- summary(neg_bin_race$fit)

coef(nb_sum)

#storing values in table and computing confidence interval
nb_table <- as.tibble(coef(nb_sum)) %>% 
  mutate(low = `Estimate` - 1.96 * `Std. Error`,
         high = `Estimate` + 1.96 * `Std. Error`) %>% 
  mutate(across(!contains("z"), exp),
         value = c("Intercept",
                   "Race - Black", 
                   "PC1", 
                   "PC2")) %>% 
  relocate(value, .before = Estimate) %>%
  relocate(c(low, high), .before = `z value`) %>%  
  mutate(across(c("low", "high"), round, 2),
          across(where(is.numeric), ~format(., scientific = FALSE))) %>% 
  unite("95% CI", c(low, high), sep = ", ", na.rm = TRUE)

##pull fitted values to save in main df
df$nb_fit <- neg_bin_race$fit$fitted.values

## pull AIC
aic_nb <- AIC(neg_bin_race$fit)

print(confint(neg_bin_race$fit))

```
Testing for spatial association. 
```{r}
#setting up matrix
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

#creating neighbor list 
  nablist_race_df <- mat2listw(h_race_df, style = "W")

# Using testing for spatial association using Moran's test (global)
  print(moran.test(df$MRR, nablist_race_df, zero.policy = TRUE, randomisation = FALSE))
# plot Moran's I
  moran.plot(df$MRR, nablist_race_df, zero.policy = TRUE)
  #save plot
png(here("results/moran_mrr.png"))
  moran.plot(df$MRR, nablist_race_df, zero.policy = TRUE)
dev.off()

  # calculate and print LISA statistics (local)
lisa_race <- localmoran(df$MRR, nablist_race_df, zero.policy=TRUE)
summary(lisa_race[,1]) 
```

#Fitting CAR model (requires neighbors)
```{r}
  h_race_df <- matrix(0, nrow(df), nrow(df))
  h_race_df[cbind(match(us_nabs$StateCode, df$`State Abbreviation`),
                  match(us_nabs$NeighborStateCode, df$`State Abbreviation`))] <- 1

  nablist_race_df <- mat2listw(h_race_df, style = "W")

  car_race <- spautolm(MRR ~   prin_comp1 + prin_comp2, 
                           data = df,
                           listw = nablist_race_df,
                           family = "CAR")

car_sum <-  summary(car_race)
car_sum

##pull fitted values to save in main df
df$car_fit <- car_race$fit$fitted.values

#Pulling estimates and standard errors
car_table <- as.tibble(coef(car_sum)) %>% 
  mutate(low = round(`Estimate` - 1.96 * `Std. Error`, 2),
         high = round(`Estimate` + 1.96 * `Std. Error`, 2),
        value = c("Intercept",
                   "PC1",
                   "PC2")) %>% 
  relocate(value, .before = Estimate) %>%
  relocate(c(low, high), .before = `z value`) %>% 
  unite("95% CI", c("low", "high"), sep = ", ", na.rm = TRUE)

#lambda confidence interval
  car_race$lambda.se
  car_race$lambda + c(-2,2) * car_race$lambda.se

##pulling AIC from CAR model
aic_car <- AIC(car_race)

```

Organizing mod_AIC into new dataframe
```{r}
aic_df <- data.frame(AIC = c(aic_mult, aic_nb, aic_car),
                     Model = c("Multiple Linear Regression", "Negative Binomial", "Conditional Autoregressive"))

aic_df <- aic_df %>%
  mutate(AIC  = round(AIC, 2))

```

```{r}
##Making kable table of model results
library(kableExtra)
library(knitr)
table_fit <- aic_df %>%
  dplyr::select(c(Model, AIC)) %>% 
  kable(caption = "<center>Model Comparison</center>", col.names = c("Model Type", "AIC"),
        align = "c") %>%
  kable_minimal() %>%
  kable_styling(font_size = 50) %>%
gsub("font-size: initial !important;",
"font-size: 50pt !important;",
.)  %>% 
  landscape()

table_fit %>%  
  save_kable(here("results/table_fit.png"))

```

```{r}
##Making kable table of summary results

library(kableExtra)

mult_table %>%
  mutate_if(is.numeric, round, 2) %>%
  mutate(`Pr(>|t|)` = ifelse(`Pr(>|t|)` < 0.005, 
                             cell_spec(paste0("<", formatC(0.005, format = "f", digits = 3)), "latex"),
                `Pr(>|t|)`)) %>%
  kable(caption = "<center>Summary of Multiple Linear Regression Model</center>",
        align = "c",
        col.names = c("Coefficient", "Estimate", "Std. Error", "95% CI", "t value", "p value"), digits = 3) %>%
  kable_minimal() %>%
  kable_styling(font_size = 50) %>%
  gsub("font-size: initial !important;",
        "font-size: 50pt !important;",
        .)  %>% 
  landscape() %>%  
  save_kable(here("results/mult_table.png"))

nb_table %>% 
 # mutate(across(where(is.numeric), ~format(., scientific = FALSE))) %>%
  mutate(across(c(2:3,5:6), as.numeric)) %>%
  mutate(`Pr(>|z|)` = round(`Pr(>|z|)`, 2),
    `Pr(>|z|)` = ifelse(`Pr(>|z|)` < 0.005, 
                             cell_spec(paste0("<", formatC(0.005, format = "f", digits = 3)), "latex"),
                `Pr(>|z|)`)) %>%
  kable(caption = "<center>Summary of Negative Binomial Regression Model</center>",
        align = "c",
        col.names = c("Coefficient", "Estimate", "Std. Error", "95% CI", "z value", "p value"), digits = 2) %>%
  kable_minimal() %>%
  kable_styling(font_size = 50) %>%
  gsub("font-size: initial !important;",
        "font-size: 50pt !important;",
        .)  %>% 
  landscape() %>%  
  save_kable(here("results/nb_table.png"))

car_table %>%
  mutate_if(is.numeric, round, 2) %>%
  mutate(`Pr(>|z|)` = ifelse(`Pr(>|z|)` < 0.005, 
                             cell_spec(paste0("<", formatC(0.005, format = "f", digits = 3)), "latex"),
                `Pr(>|z|)`)) %>%
  kable(caption = "<center>Summary of Conditional Autoregressive Model</center>",
        align = "c",
        col.names = c("Coefficient", "Estimate", "Std. Error", "95% CI", "z value", "p value"), digits = 3) %>%
  kable_minimal() %>%
  kable_styling(font_size = 50) %>%
  gsub("font-size: initial !important;",
        "font-size: 50pt !important;",
        .)  %>% 
  landscape() %>%  
  save_kable(here("results/car_table.png"))

```

```{r}
saveRDS(df, here("data/processed_data/df_fit_alt.rds"))
```