### 2023-05-05 readme.qmd

---
title: "Read Me"
format: html
editor: visual
---

This respository contains data and code to reproduce my project. Raw data is store in the `data` subfolder `raw_data`, while processed data produced by wrangling is stored under `processed_data`.

Steps to reproduce this project are:

1.  Open the R project `dataanalysis-template`
2.  Open the folder `code` and start by running files in `processing_code`. The files are numbered sequentially, but `1_cleaning_2020` is the only file in this folder that must be run in order. The others may be run in any order.
3.  Next, run the files in `analysis_code` . The files are numbered sequentially in the order that they should be run.
4.  Data files produced by these files are store in the `data` subfolder `processed_data`
5.  Images and tables produced by these files are stored in the `results` folder.
6.   The manuscript is stored in subfolder `manuscript` within the `products` folder.